### 2021-11-10 code/processing_code/processing-declarations.Rmd
Disasters Revision
Giving up on the time series data

---
title: "Data Processing: Disaster Declaration Summaries"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---
<br>

---

## Introduction
This markdown imports the raw data and performs some cleaning as well as feature engineering. The raw data comes from [the OpenFEMA Dataset website](https://www.fema.gov/openfema-data-page/disaster-declarations-summaries-v2).

<br>

Overall, processing this dataset involves the following elements:

* Raw data summarization and overview
* Removal of irrelevant variables
* Summarization of key variables for each disaster declaration and state
* Feature engineering to create duration of incident and response
* Saving the processed dataset as an .rds

<br>

To keep track of processing changes, each iteration of processed data will be named `missions_#`.

<br>

Ultimately, this processed dataset will be joined to the Disaster Declaration Summaries processed dataset for the analysis.

---

## Required Packages
The following R packages are required for this script:

* here: for path setting
* tidyverse: for all packages in the Tidyverse (ggplot2, dyplr, tidyr, readr, purr, tibble, stringr, forcats)
* skimr: for data summarization
```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #to set paths
library(tidyverse) #for data processing
library(skimr) #for data summarization
```

---

## Load Raw Data
Load the raw data downloaded from provided hyerlink.
```{r load data}
#path to data
#note the use of the here() package and not absolute paths
data_location_dss <- here::here("data", "raw_data", "DisasterDeclarationsSummaries.csv")

#load data
declarations_raw <- utils::read.csv(data_location_dss, na.strings = c("", "NA"))
```

---

## Raw Data Overview
Examine the summary and structure of the raw data.
```{r raw data overview}
#summary using base R
summary(declarations_raw)

#summary and structure using the skimr package
skimr::skim(declarations_raw)
```