### 2021-11-11 code/analysis_code/analysis-eda.Rmd
Finished processing disasters, started EDA

---
title: "Analysis: Exploratory Data Analysis"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---
<br>

---

## Introduction
This markdown imports the processed data and conducts the exploratory data analysis. Follow the processing markdowns in the `processing_code` folder to generate the processed data.

<br>

This analysis examines the predictors of allocation of FEMA funds during disaster declarations. In terms of the datasets, the two outcomes of interest are Requested Amount (`ReqAmt`) and Obligated Amount (`OblAmt`).

<br>

For each variable, the following steps will be completed:

1. Produce and print some numerical output (e.g. table, summary statistics)
2. Create a histogram or density plot (continuous variables only)
3. Scatterplot, boxplot, or other similar plots against the main outcome of interest
4. Any other exploration steps that may be useful.

The steps will be numbered for each variable as specified above.

---

## Required Packages
The following R packages are required for this script:

* here: for path setting
* tidyverse: for all packages in the Tidyverse (ggplot2, dyplr, tidyr, readr, purr, tibble, stringr, forcats)
* skimr: for data summarization
* summarytools: for overall dataframe summary
* ggplot2: for plotting data
* car: for creating QQ plots
* gt: for creating tables for summary statistics / other numerical outputs
* scales: for percent calculation

```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #to set paths
library(tidyverse) #for data processing
library(skimr) #for data summarization
library(summarytools) #for overall dataframe summary
library(ggplot2) #for plotting data
library(car) #for creating QQ plots
library(gt) #for creating tables
library(scales) #for percent calculation

#global environment options
# formatting for script to avoid scientific notation output
options(scipen=999)

# globally set theme for ggplots
ggplot2::theme_set(theme_classic())
```

---

## Load Processed Data
Load the data generated from the markdowns in the `processing_code` folder.
```{r load data}
#define file path
data_location <- here::here("data", "processed_data", "combinedprocessed.rds")

#load data
EDAdata <- readRDS(data_location)
```

---

## Data Overview
To better understand the data, let's use `summarytools` to better visualize the data.
```{r data overview}
summarytools::dfSummary(EDAdata)
```

---

## Outcome of Interest: Requested FEMA Funds
The first of two primary outcomes of interest is the total amount of funds the state requested from FEMA for a disaster declaration.
```{r}
#summary statistics
base::summary(EDAdata$ReqAmt)

#density plot
ggplot2::ggplot(data = EDAdata, aes(x = ReqAmt)) +
  geom_density()

#boxplot
ggplot2::ggplot(data = EDAdata, aes(y = ReqAmt)) +
  geom_boxplot()

#crop to $1M
ggplot2::ggplot(data = EDAdata, aes(y = ReqAmt)) +
  geom_boxplot() +
  ylim(0, 1000000)

##QQ plot
car::qqPlot(EDAdata$ReqAmt)
```

## Outcome of Interest: Obligated FEMA Funds