### 2021-11-11 code/processing_code/processing-combine.Rmd

---
title: "Data Processing: Creating the Combined Dataframe"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---
<br>

---

## Introduction
This markdown imports the processed data from disaster declarations and mission assignments and combines them into one dataframe for analysis.

---

## Required Packages
The following R packages are required for this script:

* here: for path setting
* tidyverse: for all packages in the Tidyverse (ggplot2, dyplr, tidyr, readr, purr, tibble, stringr, forcats)
```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #to set paths
library(tidyverse) #for data processing
```

---

## Load Data
Load the processed data from the `processed_data` folder.
```{r load data}
#define path to processed population data
data_location1 <- here::here("data", "processed_data", "declarationsprocessed.rds")
data_location2 <- here::here("data", "processed_data", "missionsprocessed.rds")
data_location3 <- here::here("data", "raw_data", "DisasterDeclarationsSummaries.csv")

#load data
declarations <- readRDS(data_location1)
missions <- readRDS(data_location2)
raw_decs <- utils::read.csv(data_location3, na.strings = c("", "NA"))
```

---

## Individual Dataset Overview
Examine the summary and structure of the individual datasets.
```{r dataset overview}
#summary using base R
summary(declarations)
summary(missions)
```

---

## Joining the Datasets
The two datasets could be joined by state or declaration number. 
```{r}
# #first change name of states column in missions df
# names(missions)[names(missions) == 'stateorTribe'] <- 'state'
# 
# #create df with states and declaration numbers from each df
# missions_nums <- as.data.frame(table(missions$state, missions$disasterNumber))
# declarations_nums <- as.data.frame(table(declarations$state, declarations$disasterNumber))

#we want a full join to keep all of the data
combined <- dplyr::full_join(declarations, missions, by = 'disasterNumber')
```

---

It's not a flawless join, so let's do some investigation to see where there might be gaps.
```{r}
summary(combined)

#filter the rows that aren't in declarations data but are in missions data
missing_nums <- combined %>% 
                  filter(is.na(state)) %>%
                  select(disasterNumber)

#unique list of missing declarations
missing_declarations <- unique(missing_nums)

#check missing declarations in raw declarations
decs_check <- raw_decs %>%
                dplyr::filter(disasterNumber %in% missing_declarations$disasterNumber)

#the missing values represent declarations in territories, not states
#for now, remove - we can always come back and adjust as needed
combined_processed <- combined %>%
                        dplyr::filter(!is.na(IncidentYear))
```

---

## Save Processed Data

Let's save it into the `processed_data` folder.
```{r save processed data}

#location to save processed file
combined_data_location <- here::here("data","processed_data","combinedprocessed.rds")

#save as an RDS
saveRDS(combined_processed, file = combined_data_location)
```