### 2021-11-11 code/processing_code/processing-populations.Rmd

---
title: "Data Processing: Population Estimates"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---
<br>

---

## Introduction
This markdown imports the raw data and performs some cleaning as well as feature engineering. The raw data comes from  [the U.S> Department of Agriculture Economic Research Service](https://www.ers.usda.gov/data-products/county-level-data-sets/download-data/).

<br>

Overall, processing this dataset involves the following elements:

* Raw data summarization and overview
* Removal of irrelevant variables
* Feature engineering to determine number of counties per state
* Saving the processed dataset as an .rds

<br>

To keep track of processing changes, each iteration of processed data will be named `pops_#`. The final processed dataset will be `populations_processed`.

<br>

Ultimately, this processed dataset will be joined to the Disaster Declaration Summaries processed dataset for the analysis.

---

## Required Packages
The following R packages are required for this script:

* here: for path setting
* tidyverse: for all packages in the Tidyverse (ggplot2, dyplr, tidyr, readr, purr, tibble, stringr, forcats)
* skimr: for data summarization
```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #to set paths
library(tidyverse) #for data processing
library(skimr) #for data summarization
library(lubridate) #for date wrangling
```

---

## Load Raw Data
Load the raw data downloaded from provided hyerlink.
```{r load data}
#path to data
#note the use of the here() package and not absolute paths
data_location_pops <- here::here("data", "raw_data", "PopulationEstimates.csv")

#load data
populations_raw <- utils::read.csv(data_location_pops, na.strings = c("", "NA"))
```

---

## Raw Data Overview
Examine the summary and structure of the raw data.
```{r raw data overview}
#summary using base R
summary(populations_raw)

#summary and structure using the skimr package
skimr::skim(populations_raw)
```
There are 14 variables and 3,283 observations. The data is structured in such a way that it includes totals for the states as well as for each county - time to clean!

---

## Data Cleaning and Wrangling
First, let's remove the unimportant variables.
```{r variable list}
#print list of variables
colnames(populations_raw)
```

We only want to keep the following variables:

* FIPStxt
* State
* Area.name
* Population.2020
```{r variable selection}
#select key variables
pops_1 <- populations_raw %>%
                dplyr::select(FIPStxt,
                              State,
                              Area.name,
                              Population.2020)
```
We will eventually get rid of FIPStxt, but we need it to separate state and county level data.

---

To create the state-level subset, filter the rows based on values of FIPStxt.
```{r state filter}
#create a vector of FIPStxt values
state_fipstxt  <- c("01000",
                    "02000",
                    "03000",
                    "04000",
                    "05000",
                    "06000",
                    "07000",
                    "08000",
                    "09000",
                    "10000",
                    "11000",
                    "12000",
                    "13000",
                    "14000",
                    "15000",
                    "16000",
                    "17000",
                    "18000",
                    "19000",
                    "20000",
                    "21000",
                    "22000",
                    "23000",
                    "25000",
                    "26000",
                    "27000",
                    "28000",
                    "29000",
                    "30000",
                    "31000",
                    "32000",
                    "33000",
                    "34000",
                    "35000",
                    "36000",
                    "37000",
                    "38000",
                    "39000",
                    "40000",
                    "41000",
                    "42000",
                    "43000",
                    "44000",
                    "45000",
                    "46000",
                    "47000",
                    "48000",
                    "49000",
                    "50000",
                    "51000",
                    "52000",
                    "53000",
                    "54000",
                    "55000",
                    "56000",
                    "60000",
                    "66000",
                    "69000",
                    "72000",
                    "78000")

#subset based on which ones are in this vector
#then drop FIPStxt and Area.name
pops_states <- pops_1 %>%
                dplyr::filter(FIPStxt %in% state_fipstxt) %>%
                subset(select = -c(FIPStxt,
                                   Area.name))
```
Now we have a dataframe with state abbreviations and population estimates.

---

The last step is to count the number of counties per state and add it to the state dataframe.
```{r counties count}
#subset pops_1 data that doesn't include states
#remove first row too since it's the entire US
pops_2 <- pops_1 %>%
            dplyr::filter(! FIPStxt %in% state_fipstxt) %>%
            dplyr::slice(-1)

#count the number of occurrences of each state (which represents the number of counties)
pops_3 <- as.data.frame(table(pops_2$State))

#relabel columns
names(pops_3) <- c("State", "Counties")

#combine with states dataframe
pops_processed <- merge(pops_states, pops_3, by="State")
```

---

## Processed Data Overview
We have our processed dataset! Let's take a look at the overview to see how it changed.
```{r processed data overview}
#summary using base R
summary(pops_processed)

#summary and structure using the skimr package
skimr::skim(pops_processed)
```

---

## Save Processed Data

We now have our processed dataset for state populations and counties! Let's save it into the `processed_data` folder.
```{r save processed data}

#location to save processed file
pops_data_location <- here::here("data","processed_data","popsprocessed.rds")

#save as an RDS
saveRDS(pops_processed, file = pops_data_location)
```