### 2021-11-11 code/analysis_code/analysis-linearmodels.Rmd

---
title: "Analysis: Linear Models"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---
<br>

---

## Introduction
This markdown imports the processed analysis data and uses the tidymodels framework to fit linear models. Follow the previous markdowns to generate the data.

<br>

This analysis examines the predictors of allocation of FEMA funds during disaster declarations. The following definitions exist for the analysis:

* The two outcomes of interest are Requested Amount (`ReqAmt`) and Obligated Amount (`OblAmt`).
* The primary predictor of interest is Incident Duration (`IncidentDuration`)
* Whichever outcome is not currently fitted to the model will *not* be considered a predictor of interest

<br>

For each outcome of interest, the following steps will be completed:

1. Fit a simple linear regression with the primary predictor of interest
2. Fit a multivariable linear regression
3. Fit a LASSO linear regression

---

## Required Packages
The following R packages are required for this script:

* here: for path setting
* tidyverse: for all packages in the Tidyverse (ggplot2, dyplr, tidyr, readr, purr, tibble, stringr, forcats)
* skimr: for data summarization
* tidymodels: for data modeling

```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #to set paths
library(tidyverse) #for data processing
library(skimr) #for data summarization
library(tidymodels) #for data modeling

#global environment options
# formatting for script to avoid scientific notation output
options(scipen=999)
```

---

## Load Processed Data
Load the data generated from the markdowns in the `processing_code` folder.
```{r load data}
#define file path
data_location <- here::here("data", "processed_data", "analysisdata.rds")

#load data
analysis_data <- readRDS(data_location)
```

---

## Data Overview
To better understand the data, let's use `summarytools` to better visualize the data.
```{r data overview}
skimr::skim(analysis_data)
```

---

## Data Filtering
We will make two datasets with the only the relevant variables for each:

* state
* IncidentYear
* IncidentMonth
* declarationType
* incidentType
* IncidentDuration
* IH
* PA
* HM
* Population
* Counties
* TotalAgencies
* FEMACSAvg

```{r data subsets}
#obligated amount
obl_data <- analysis_data %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, OblAmt)

#requested amount
req_data <- analysis_data %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, ReqAmt)
```

---

## Outcome of Interest: Requested FEMA Funds
We will start with requested FEMA funds as the outcome of interest.

<br>

### Data Set Up
Split the data randomly into train and test datasets, and define cross validation
```{r req data splitting}
#fix random numbers by setting the seed (helps reproducibility)
set.seed(123)

#put 3/4 of data into the training set
req_data_split <- rsample::initial_split(req_data, prop = 3/4)

#create dataframes for the two sets
req_train_data <- rsample::training(req_data_split)
req_test_data <- rsample::testing(req_data_split)
```

---

### Simple Linear Regression
Create the SLR using the `tidymodels` setup.
```{r req SLR}
#define lr mod
lr_mod <- parsnip::linear_reg() %>%
          parsnip::set_engine("lm")

#define recipe
req_SLR_rec <- recipes::recipe(ReqAmt ~ IncidentDuration, data = req_train_data)

#define SLR workflow
req_SLR_wflow <- workflows::workflow() %>%
                   workflows::add_model(lr_mod) %>%
                   workflows::add_recipe(req_SLR_rec)

#fit the training data to the workflow
req_SLR_fit <- req_SLR_wflow %>%
                  parsnip::fit(data = req_train_data)

#create tibble for model fit using broom and extract
req_SLR_fit %>%
  workflowsets::extract_fit_parsnip() %>%
  broom::tidy()
```

<br>

Create box and whisker plot for the fit output.
```{r}
req_SLR_bp <- broom.mixed::tidy(req_SLR_fit) %>%
                dotwhisker::dwplot(dot_args = list(size = 2, color = "blue"),
                                   whisker_args = list(color = "blue"),
                                   vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
req_SLR_bp
```

<br>

Predict requested funds from FEMA  based on the model above for the training data set.
```{r req SLR predict}
#use predict to predict funds
stats::predict(req_SLR_fit, req_train_data)

#create df with model predictions and actual measures
req_SLR_aug <- augment(req_SLR_fit, req_train_data)

#add labeling variables for later summarization
req_SLR_train <- req_SLR_aug %>%
                    dplyr::mutate(data = "train",
                                  model = "SLR")
```

<br>

Assess goodness of fit measures.
```{r req SLR eval}
#using glace
modelsummary::glance(req_SLR_fit)

#using the prediction capacity
req_SLR_train %>% yardstick::metrics(truth = ReqAmt, estimate = .pred)
```
This is a terrible model. Time to try adding more predictors.

---

### Multivariable Linear Regression
Create the MVR using the `tidymodels` setup.
```{r req SLR}
#define linear mod
lr_mod <- linear_reg() %>%
              set_engine("lm")

#define recipe
req_MVR_rec <- recipes::recipe(ReqAmt ~ ., data = req_train_data)

#define SLR workflow
req_MVR_wflow <- workflows::workflow() %>%
                   workflows::add_model(lr_mod) %>%
                   workflows::add_recipe(req_MVR_rec)

#fit the training data to the workflow
req_MVR_fit <- req_MVR_wflow %>%
                  parsnip::fit(data = req_train_data)

#create tibble for model fit using broom and extract
req_MVR_fit %>%
  workflowsets::extract_fit_parsnip() %>%
  broom::tidy()
```

<br>

Create box and whisker plot for the fit output.
```{r}
#basic box and whisker plot
req_MVR_bp <- broom.mixed::tidy(req_MVR_fit) %>%
                dotwhisker::dwplot(dot_args = list(size = 2, color = "blue"),
                                   whisker_args = list(color = "blue"),
                                   vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
req_MVR_bp

#there's a lot of information here, but hard to identify the ones that are significant due to volume

#box and whisker plot for significant predictors
#first filter significant results
req_MVR_fit_sig <- broom.mixed::tidy(req_MVR_fit) %>%
                      dplyr::filter(p.value < 0.05)

#box and whisker plot for lm_fit2 significant predictors
req_MVR_fit_bp2 <- req_MVR_fit_sig %>%
                        dotwhisker::dwplot(dot_args = list(size = 2, color = "blue"),
                                           whisker_args = list(color = "blue"),
                                           vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
req_MVR_fit_bp2
```

<br>

Predict requested funds from FEMA  based on the model above for the training data set.
```{r req SLR predict}
#use predict to predict funds
stats::predict(req_MVR_fit, req_train_data)

#create df with model predictions and actual measures
req_MVR_aug <- augment(req_MVR_fit, req_train_data)

#add labeling variables for later summarization
req_MVR_train <- req_MVR_aug %>%
                    dplyr::mutate(data = "train",
                                  model = "MVR")
```

<br>

Assess goodness of fit measures.
```{r req SLR eval}
#using glace
modelsummary::glance(req_SLR_fit)

#using the prediction capacity
req_SLR_train %>% yardstick::metrics(truth = ReqAmt, estimate = .pred)
```
This is a terrible model. Time to try adding more predictors.