### 2021-11-11 code/processing_code/processing-combine.Rmd

---
title: "Data Processing: Creating the Combined Dataframe"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---
<br>

---

## Introduction
This markdown imports the processed data from disaster declarations and mission assignments and combines them into one dataframe for analysis.

---

## Required Packages
The following R packages are required for this script:

* here: for path setting
* tidyverse: for all packages in the Tidyverse (ggplot2, dyplr, tidyr, readr, purr, tibble, stringr, forcats)
```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #to set paths
library(tidyverse) #for data processing
```

---

## Load Data
Load the processed data from the `processed_data` folder.
```{r load data}
#define path to processed population data
data_location1 <- here::here("data", "processed_data", "declarationsprocessed.rds")
data_location2 <- here::here("data", "processed_data", "missionsprocessed.rds")
data_location3 <- here::here("data", "raw_data", "DisasterDeclarationsSummaries.csv")

#load data
declarations <- readRDS(data_location1)
missions <- readRDS(data_location2)
raw_decs <- utils::read.csv(data_location3, na.strings = c("", "NA"))
```

---

## Individual Dataset Overview
Examine the summary and structure of the individual datasets.
```{r dataset overview}
#summary using base R
summary(declarations)
summary(missions)
```

---

## Preparing the datasets
The two datasets could be joined by state or declaration number. 
```{r}
# first change name of states column in missions df
names(missions)[names(missions) == 'stateorTribe'] <- 'state'

#find differences between disasterNumber and states included in each dataset
declarations_sub <- declarations %>% dplyr::select(disasterNumber,
                                                   state)

missions_sub <- missions %>% dplyr::select(disasterNumber,
                                           state)

missing_declarations <- dplyr::setdiff(missions_sub, declarations_sub)

#search in raw declarations
decs_check <- raw_decs %>%
                dplyr::filter(disasterNumber %in% missing_declarations$disasterNumber)
```
By examining the `decs_check` df, we can see that the missing declarations from the missions assignments do not exist in the declarations summaries. We can move forward and accept the reduction in data.

---

It's not a flawless join, so let's do some investigation to see where there might be gaps.
```{r}
#use a left join to merge into the declarations df
combined <- dplyr::left_join(declarations, missions, by = c("state" = "state", "disasterNumber" = "disasterNumber"))

#the NAs for funding cost are declarations for which FEMA did not provide any funding
#therefore they can be replaced with 0
combined$TotalAgencies[is.na(combined$TotalAgencies)] <- 0
combined$ReqAmt[is.na(combined$ReqAmt)] <- 0
combined$OblAmt[is.na(combined$OblAmt)] <- 0
combined$FEMACSAvg[is.na(combined$FEMACSAvg)] <- 0

#summary of processed data
summary(combined)

```

---

## Save Processed Data

Let's save it into the `processed_data` folder.
```{r save processed data}

#location to save processed file
combined_data_location <- here::here("data","processed_data","combinedprocessed.rds")

#save as an RDS
saveRDS(combined, file = combined_data_location)
```