### 2021-11-12 code/analysis_code/analysis-ML-RA.Rmd
Finished linear models, started ML models

---
title: "Analysis: Machine Learning - ReqAmt"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---

---

## Introduction

This script uses requested FEMA funds as the outcome of interest and fits the following models to the analysis data:

* Null
* Decision Tree
* LASSO
* Random Forest.

It compares the  models, and then finally fits the “best” model to the test data.

<br>

Follow the previous processing and analysis markdowns to generate the analysis data used here.

<br>

Each model will follow this process:

1. Model Specification
2. Workflow Definition
3. Tuning Grid Specification
4. Tuning Using Cross-Validation and the tune_grid() function
5. Identify Best Model
6. Model Evaluation

---

## Required Packages
The following R packages are required for this script:

* here: for data loading/saving
* tidyverse: for data management
* tidymodels: for data modeling
* skimr: for variable summaries
* broom.mixed: for converting bayesian models to tidy tibbles
* rpart.plot: for visualizing a decision tree
* vip: for variable importance plots
* glmnet: for lasso models
* doParallel: for parallel backend for tuning processes
* ranger: for random forest models
```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(rpart.plot) #for visualizing a decision tree
library(vip) #for variable importance plots
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models

#global environment options
# formatting for script to avoid scientific notation output
options(scipen=999)
```

---

## Load Data
Load the analysis data from the `processed_data` folder in the project file.
```{r load data}
#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","analysisdata.rds")

#load data. 
analysisdata <- readRDS(data_location)

#summary of data using skimr package
skimr::skim(analysisdata)
```

---

## Data Setup
We will specify the following parameters:

* Drop the unused variables
* Transform Incident Month into ordinal variable
* Set the random seed to 123
* Split the dataset into 70% training, 30% testing, stratified on ReqAmt
* 5-fold cross validation, 5 times repeated, stratified on ReqAmt for the CV folds
* Create a recipe for data and fitting that codes categorical variables as dummy variables
```{r data setup}
#requested amount filter
req_data <- analysisdata %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, ResponseDays, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, ReqAmt)

#transform incident month into ordinal variable
req_data$IncidentMonth <- ordered(req_data$IncidentMonth, levels = 1:12,
                                  labels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))

#set random seed
set.seed(123)

#split dataset into 70% training, 30% testing
#use ReqAmt as stratification
req_data_split <- rsample::initial_split(req_data, prop = 7/10,
                                     strata = c(ReqAmt, state, incidentType))

#optional tag row for later identification: 
req_data$rowid<-1:nrow(req_data)
dplyr_stratified <- function(df, percent, ...){
  columns<-enquos(...)
   #group then sample each group
  out<-df %>% group_by(!!!columns)  %>% slice( sample(1:n(), percent*n())) 
}

train_group <- dplyr_stratified(req_data, 0.9, state, incidentType)
train_group

test_group <- dplyr_stratified(req_data, 0.5, state, incidentType)
test_group

#create dataframes for the two sets:
req_train_data <- rsample::training(req_data_split)
req_test_data <- rsample::testing(req_data_split)

#training set proportions by ReqAmt
req_train_data %>%
      dplyr::count(ReqAmt) %>%
      dplyr::mutate(prop = n / sum(n))

#testing set proportions by ReqAmt
req_test_data %>%
      dplyr::count(ReqAmt) %>%
      dplyr::mutate(prop = n / sum(n))

#5-fold cross validation, 5 times repeated, stratified on `ReqAmt` for the CV folds
req_folds <- rsample::vfold_cv(req_train_data,
                               v = 5,
                               repeats = 5,
                               strata = ReqAmt)

#create recipe that codes categorical variables as dummy variables
RA_rec <- recipes::recipe(ReqAmt ~ ., data = req_train_data) %>%
          recipes::step_dummy(all_nominal_predictors())
```

---

## Null Model
Determine the performance of a null model (i.e. one with no predictors) and compute the RMSE for both training and test data.
```{r null model}
#create null model
null_mod <- parsnip::null_model() %>%
            parsnip::set_engine("parsnip") %>%
            parsnip::set_mode("regression")

#add recipe and model into workflow
null_wflow <- workflows::workflow() %>%
              workflows::add_recipe(RA_rec) %>%
              workflows::add_model(null_mod)

#"fit" model to training data
req_null_train <- null_wflow %>%
                  parsnip::fit(data = req_train_data)

#summary of null model with training data to get mean (which in this case is the RMSE)
req_null_train_sum <- broom.mixed::tidy(req_null_train)
req_null_train_sum

#RMSE for training data for formatting later
req_null_RMSE_train <- tibble::tibble(
                          rmse = rmse_vec(truth = req_train_data$ReqAmt,
                                              estimate = rep(mean(req_train_data$ReqAmt), nrow(req_train_data))),
                          SE = 0,
                          model = "Null - Train")

#"fit" model to test data
req_null_test <- null_wflow %>%
                  parsnip::fit(data = req_test_data)

#summary of null model with test data to get mean (which in this case is the RMSE)
req_null_test_sum <- broom.mixed::tidy(req_null_test)
req_null_test_sum

#RMSE for testing data for formatting later
req_null_RMSE_test <- tibble::tibble(
                          rmse = rmse_vec(truth = req_test_data$ReqAmt,
                                              estimate = rep(mean(req_test_data$ReqAmt), nrow(req_test_data))),
                          SE = 0,
                          model = "Null - Test")
```

---

## LASSO Model

<br>

### 1. Model Specification
```{r lasso model specification}
#define the lasso model
#mixture = 1 identifies the model to be a LASSO model
lasso_mod <-
  parsnip::linear_reg(mode = "regression",
                      penalty = tune(), 
                      mixture = 1) %>%
  parsnip::set_engine("glmnet")

#use the recipe specified earlier
```

<br>

### 2. Workflow Definition
```{r lasso wflow def}
#define workflow for lasso regression
lasso_wflow <- workflows::workflow() %>%
               workflows::add_model(lasso_mod) %>%
               workflows::add_recipe(RA_rec)
```

<br>

### 3. Tuning Grid Specification
```{r lasso tune grid spec}
#tuning grid specification
lasso_grid <- grid_regular(penalty(), levels = 10)

#5 lowest penalty values
lasso_grid %>%
  dplyr::top_n(-5)

#5 highest penalty values
lasso_grid %>%
  dplyr::top_n(5)
```

<br>

### 4. Tuning Using Cross-Validation and the `tune_grid()` function 
```{r lasso tuning}
#tune the model with previously specified cross-validation and RMSE as target metric
lasso_res <- lasso_wflow %>%
                tune::tune_grid(resamples = req_folds,
                                grid = lasso_grid,
                                control = control_grid(verbose = TRUE, 
                                                       save_pred = TRUE),
                                metrics = metric_set(rmse))

#look at 15 models with lowest RMSEs
top_lasso_models <- lasso_res %>%
                      tune::show_best("rmse", n = 15) %>%
                      dplyr::arrange(penalty)
top_lasso_models

#default visualization
lasso_res %>% autoplot()
```