### 2021-11-12 code/analysis_code/analysis-ML-OA.Rmd

---
title: "Analysis: Machine Learning - OblAmt"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---

---

## Introduction

This script uses obligated FEMA funds as the outcome of interest and fits the following models to the analysis data:

* Null
* Decision Tree
* Random Forest
* Bagged Tree

It compares the  models, and then finally fits the “best” model to the test data.

<br>

Follow the previous processing and analysis markdowns to generate the analysis data used here.

<br>

Each model will follow this process:

1. Model Specification
2. Workflow Definition
3. Tuning Grid Specification
4. Tuning Using Cross-Validation and the tune_grid() function
5. Identify Best Model
6. Model Evaluation

---

## Required Packages
The following R packages are required for this script:

* here: for data loading/saving
* tidyverse: for data management
* tidymodels: for data modeling
* skimr: for variable summaries
* broom.mixed: for converting bayesian models to tidy tibbles
* rpart.plot: for visualizing a decision tree
* vip: for variable importance plots
* glmnet: for lasso models
* doParallel: for parallel backend for tuning processes
* ranger: for random forest models
* baguette: for bagged decision tree
```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(rpart.plot) #for visualizing a decision tree
library(vip) #for variable importance plots
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(baguette) #for bagged decision tree

#global environment options
# formatting for script to avoid scientific notation output
options(scipen=999)
```

---

## Load Data
Load the analysis data from the `processed_data` folder in the project file.
```{r load data}
#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","analysisdata.rds")

#load data. 
analysisdata <- readRDS(data_location)

#summary of data using skimr package
skimr::skim(analysisdata)
```

---

## Data Setup
We will specify the following parameters:

* Drop the unused variables
* Perform feature engineering
* Transform Incident Month into ordinal variable
* Set the random seed to 123
* Split the dataset into 70% training, 30% testing, stratified on incidentType
* 5-fold cross validation, 5 times repeated, stratified on incidentType for the CV folds
* Create a recipe for data and fitting that codes categorical variables as dummy variables
```{r data setup}
#requested amount filter
obl_data <- analysisdata %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, ResponseDays, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, OblAmt)

#transform incident month into ordinal variable
obl_data$IncidentMonth <- ordered(obl_data$IncidentMonth, levels = 1:12,
                                  labels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))

#not enough values per state, so create a variable that is FEMA Regions
obl_data$FEMARegion <- ifelse((obl_data$state == "CT" |
                                 obl_data$state == "ME" |
                                 obl_data$state == "MA" |
                                 obl_data$state == "NH" |
                                 obl_data$state == "RI" |
                                 obl_data$state == "VT"),
                              "Region I",
                              ifelse((obl_data$state == "NJ" |
                                         obl_data$state == "NY" |
                                         obl_data$state == "PR" |
                                         obl_data$state == "VI"),
                                      "Region II",
                                       ifelse((obl_data$state == "DE" |
                                                 obl_data$state == "MD" |
                                                 obl_data$state == "PA" |
                                                 obl_data$state == "VA" |
                                                 obl_data$state == "DC" |
                                                 obl_data$state == "WV"),
                                              "Region III",
                                               ifelse((obl_data$state == "AL" |
                                                         obl_data$state == "FL" |
                                                         obl_data$state == "GA" |
                                                         obl_data$state == "KY" |
                                                         obl_data$state == "MS" |
                                                         obl_data$state == "NC" |
                                                         obl_data$state == "SC" |
                                                         obl_data$state == "TN"),
                                                      "Region IV",
                                                      ifelse((obl_data$state == "IL" |
                                                               obl_data$state == "IN" |
                                                               obl_data$state == "MI" |
                                                               obl_data$state == "MN" |
                                                               obl_data$state == "OH" |
                                                               obl_data$state == "WI"),
                                                            "Region V",
                                                             ifelse((obl_data$state == "AR" |
                                                                         obl_data$state == "LA" |
                                                                         obl_data$state == "NM" |
                                                                         obl_data$state == "OK" |
                                                                         obl_data$state == "TX"),
                                                                      "Region VI",
                                                                       ifelse((obl_data$state == "IA" |
                                                                                 obl_data$state == "KS" |
                                                                                 obl_data$state == "MO" |
                                                                                 obl_data$state == "NE"),
                                                                              "Region VII",
                                                                               ifelse((obl_data$state == "CO" |
                                                                                         obl_data$state == "MT" |
                                                                                         obl_data$state == "ND" |
                                                                                         obl_data$state == "SD" |
                                                                                         obl_data$state == "UT" |
                                                                                         obl_data$state == "WY"),
                                                                                      "Region VIII",
                                                                                       ifelse((obl_data$state == "AZ" |
                                                                                                 obl_data$state == "CA" |
                                                                                                 obl_data$state == "HI" |
                                                                                                 obl_data$state == "NV"),
                                                                                              "Region IX",
                                                                                               ifelse((obl_data$state == "AK" |
                                                                                                         obl_data$state == "ID" |
                                                                                                         obl_data$state == "OR" |
                                                                                                         obl_data$state == "WA"),
                                                                                                      "Region X",
                                                                                                      NA
                                                                                                      )
                                                                                              )
                                                                                      )
                                                                              )
                                                                      )
                                                            )
                                                      )
                                              )
                                     )
                              )

#drop state variable
obl_data <- obl_data %>%
              dplyr::select(-c(state))

#add to "other" category for incident Types that are infrequent
obl_data$incidentType[(obl_data$incidentType == "Chemical" |
                        obl_data$incidentType == "Coastal Storm" |
                        obl_data$incidentType == "Dam/Levee Break" |
                        obl_data$incidentType == "Mud/Landslide" |
                        obl_data$incidentType == "Volcano" |
                        obl_data$incidentType == "Terrorist" |
                        obl_data$incidentType == "Snow")] <- "Other"

#set random seed
set.seed(123)

#split dataset into 66.6% training, 33.3% testing
#use incidentType as stratification
obl_data_split <- rsample::initial_split(obl_data, prop = 2/3,
                                     strata = incidentType)

#create dataframes for the two sets:
obl_train_data <- rsample::training(obl_data_split)
obl_test_data <- rsample::testing(obl_data_split)

#5-fold cross validation, repeated 5 times, stratified by incidentType
obl_folds <- rsample::vfold_cv(obl_train_data,
                               strata = incidentType,
                               v = 5,
                               repeats = 5)

#create recipe that codes categorical variables as dummy variables
OA_rec <- recipes::recipe(OblAmt ~ ., data = obl_train_data) %>%
          recipes::step_dummy(all_nominal_predictors())
```

---

## Null Model
Determine the performance of a null model (i.e. one with no predictors) and compute the RMSE for both training and test data.
```{r null model}
#create null model
null_mod <- parsnip::null_model() %>%
            parsnip::set_engine("parsnip") %>%
            parsnip::set_mode("regression")

#add recipe and model into workflow
null_wflow <- workflows::workflow() %>%
              workflows::add_recipe(OA_rec) %>%
              workflows::add_model(null_mod)

#"fit" model to training data
obl_null_train <- null_wflow %>%
                  parsnip::fit(data = obl_train_data)

#summary of null model with training data to get mean (which in this case is the RMSE)
obl_null_train_sum <- broom.mixed::tidy(obl_null_train)
obl_null_train_sum

#RMSE for training data for formatting later
obl_null_RMSE_train <- tibble::tibble(
                          rmse = rmse_vec(truth = obl_train_data$OblAmt,
                                              estimate = rep(mean(obl_train_data$OblAmt), nrow(obl_train_data))),
                          SE = 0,
                          model = "Null - Train")

#"fit" model to test data
obl_null_test <- null_wflow %>%
                  parsnip::fit(data = obl_test_data)

#summary of null model with test data to get mean (which in this case is the RMSE)
obl_null_test_sum <- broom.mixed::tidy(obl_null_test)
obl_null_test_sum

#RMSE for testing data for formatting later
obl_null_RMSE_test <- tibble::tibble(
                          rmse = rmse_vec(truth = obl_test_data$OblAmt,
                                              estimate = rep(mean(obl_test_data$OblAmt), nrow(obl_test_data))),
                          SE = 0,
                          model = "Null - Test")
```

---

## Decision Tree

<br>

### 1. Model Specification
```{r}
#run parallels to determine number of cores
cores <- parallel::detectCores() - 1
cores

cl <- makeCluster(cores)

registerDoParallel(cl)

#define the tree model
tree_mod <-
  parsnip::decision_tree(cost_complexity = tune(),
                         tree_depth = tune(),
                         min_n = tune()) %>%
  parsnip::set_engine("rpart") %>%
  parsnip::set_mode("regression")

#use the recipe specified earlier (line 133)
```
---

### 2. Workflow Definition
```{r}
#define workflow for tree
tree_wflow <- workflows::workflow() %>%
               workflows::add_model(tree_mod) %>%
               workflows::add_recipe(OA_rec)
```

---

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
tree_grid <- dials::grid_regular(cost_complexity(),
                                 tree_depth(),
                                 min_n(),
                                 levels = 5)

#tree depth
tree_grid %>%
  dplyr::count(tree_depth)
```

---

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
tree_res <- tree_wflow %>%
                tune::tune_grid(resamples = obl_folds,
                                grid = tree_grid,
                                control = control_grid(verbose = TRUE),
                                metrics = yardstick::metric_set(rmse))

#collect metrics
tree_res %>% workflowsets::collect_metrics()

#default visualization
DT_auto_plot <- tree_res %>% autoplot()
DT_auto_plot

#save file
DT_auto_fig_file = here("results","DT-auto-plot-obl.png")
ggsave(filename = DT_auto_fig_file, plot = DT_auto_plot)

#more detailed plot
DT_detail_plot <- tree_res %>%
                    workflowsets::collect_metrics() %>%
                    dplyr::mutate(tree_depth = factor(tree_depth)) %>%
                    ggplot2::ggplot(aes(cost_complexity, mean, color = tree_depth)) +
                             geom_line(size = 1.5, alpha = 0.6) +
                             geom_point(size = 2) +
                             facet_wrap(~ .metric, scales = "free", nrow = 2) +
                             scale_x_log10(labels = scales::label_number()) +
                             scale_color_viridis_d(option = "plasma", begin = 0.9, end = 0)
DT_detail_plot

#save file
DT_detail_fig_file = here("results","DT-det-plot-obl.png")
ggsave(filename = DT_detail_fig_file, plot = DT_detail_plot)

```

---

### 5. Identify Best Model
```{r}
#select the tree model with the lowest rmse
tree_lowest_rmse <- tree_res %>%
                        tune::select_best("rmse")

#finalize the workflow by using the selected DT model
best_tree_wflow <- tree_wflow %>%
                      tune::finalize_workflow(tree_lowest_rmse)
best_tree_wflow

#one last fit on the training data
best_tree_fit <- best_tree_wflow %>%
                    parsnip::fit(data = obl_train_data)
```

---

### 6. Model evaluation
```{r}
#plot the tree
rpart.plot::rpart.plot(x = workflowsets::extract_fit_parsnip(best_tree_fit)$fit,
                                         roundint = F,
                                         type = 5,
                                         digits = 5,
                                         main = "Selected Decision Tree Model")
#save file manually -- ggsave throws an error

#find predictions and intervals
tree_resid <- best_tree_fit %>%
                  broom.mixed::augment(new_data = obl_train_data) %>%
                  dplyr::select(.pred, OblAmt) %>%
                  dplyr::mutate(.resid = .pred - OblAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
DT_pred_act <- ggplot2::ggplot(tree_resid, aes(x = .pred, y = OblAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point() +
                        labs(title = "Decision Tree Fit: Predicted vs. Actual Obligated Funds",
                              x = "Predicted Obligated Funds ($)",
                              y = "Actual Obligated Funds ($)")
DT_pred_act

#save file
DT_pred_act_file = here("results","DT-pred-act-obl.png")
ggsave(filename = DT_pred_act_file, plot = DT_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
DT_resid <- ggplot2::ggplot(tree_resid, aes(x = as.numeric(row.names(tree_resid)), y = .resid))+
                     geom_hline(yintercept = 0, color = "red", lty = 2) +
                     geom_point() +
                     labs(title = "Decision Tree Fit: Residuals",
                          x = "Observation Number",
                          y = "Residual")
DT_resid

#save file
DT_resid_file = here("results","DT-resid-obl.png")
ggsave(filename = DT_resid_file, plot = DT_resid)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
DT_pred_resid <- ggplot2::ggplot(tree_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Decision Tree Fit: Residuals vs Fitted Obligated Funds",
                                x = "Predicted Obligated Funds ($)",
                                y = "Residual")
DT_pred_resid

#save file
DT_pred_resid_file = here("results","DT-pred-resid-obl.png")
ggsave(filename = DT_pred_resid_file, plot = DT_pred_resid)

#print model performance
#print 10 best performing hyperparameter sets
tree_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err, cost_complexity) %>%
  dplyr::mutate(rmse = round(rmse, 3),
                std_err = round(std_err, 4),
                cost_complexity = scales::scientific(cost_complexity))

#print the best model performance
tree_performance <- tree_res %>% tune::show_best(n = 1)
print(tree_performance)

#compare model performance to null model
tree_RMSE <- tree_res %>%
                tune::show_best(n = 1) %>%
                dplyr::transmute(
                  rmse = round(mean, 3),
                  SE = round(std_err, 4),
                  model = "Tree") %>%
               dplyr::bind_rows(obl_null_RMSE_train)
tree_RMSE
```

<br>

---

## Random Forest Model

<br>

### 1. Model Specification
```{r}
#run parallels to determine number of cores
cores <- parallel::detectCores() - 1
cores

cl <- makeCluster(cores)

registerDoParallel(cl)

#define the RF model
RF_mod <-
  parsnip::rand_forest(mtry = tune(),
                       min_n = tune(),
                       trees = tune()) %>%
  parsnip::set_engine("ranger",
                      importance = "permutation") %>%
  parsnip::set_mode("regression")

#use the recipe specified earlier (line 133)

#check to make sure identified parameters will be tuned
RF_mod %>% tune::parameters()
```

---

### 2. Workflow Definition
```{r}
#define workflow for RF regression
RF_wflow <- workflows::workflow() %>%
               workflows::add_model(RF_mod) %>%
               workflows::add_recipe(OA_rec)
```

---

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
RF_grid <- expand.grid(mtry = c(3, 4, 5, 6),
                       min_n = c(40, 50, 60),
                       trees = c(500,1000))
```

---

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
RF_res <- RF_wflow %>%
              tune::tune_grid(resamples = obl_folds,
                              grid = RF_grid,
                              control = control_grid(verbose = TRUE, save_pred = TRUE),
                              metrics = metric_set(rmse))

#look at top 5 RF models
top_RF_models <- RF_res %>%
                    tune::show_best("rmse", n = 5)
top_RF_models

#default visualization
RF_auto_plot <- RF_res %>% autoplot()
RF_auto_plot

#save file
RF_auto_fig_file = here("results","RF-auto-plot-obl.png")
ggsave(filename = RF_auto_fig_file, plot = RF_auto_plot)
```

---

### 5. Identify Best Model
```{r}
#select the RF model with the lowest rmse
RF_lowest_rmse <- RF_res %>%
                      tune::select_best("rmse")

#finalize the workflow by using the selected RF model
best_RF_wflow <- RF_wflow %>%
                      tune::finalize_workflow(RF_lowest_rmse)
best_RF_wflow

#one last fit on the training data
best_RF_fit <- best_RF_wflow %>%
                    parsnip::fit(data = obl_train_data)
```

---

### 6. Model evaluation
```{r}
#extract model from final fit
x_RF <- best_RF_fit$fit$fit$fit

#plot most important predictors in the model
RF_plot <- vip::vip(x_RF, num_features = 20)
RF_plot

#save  manually - ggsave throws an error

#find predictions and intervals
RF_resid <- best_RF_fit %>%
                broom.mixed::augment(new_data = obl_train_data) %>%
                dplyr::select(.pred, OblAmt) %>%
                dplyr::mutate(.resid = .pred - OblAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline is a 45 degree line, along which the results should fall
RF_pred_act <- ggplot2::ggplot(RF_resid, aes(x = .pred, y = OblAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) +
                        geom_point() +
                        labs(title = "RF Fit: Actual vs. Predicted Obligated Funds",
                              x = "Predicted Obligated Funds ($)",
                              y = "Actual Obligated Funds ($)")

#save file
RF_pred_act_file = here("results","RF-pred-act-obl.png")
ggsave(filename = RF_pred_act_file, plot = RF_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
RF_resids <- ggplot2::ggplot(RF_resid, aes(x = as.numeric(row.names(RF_resid)), y = .resid))+
                            geom_hline(yintercept = 0, color = "red", lty = 2) +
                            geom_point() +
                            labs(title = "RF Fit: Residuals",
                                  x = "Observation Number",
                                  y = "Residual")
RF_resids

#save file
RF_resid_file = here("results","RF-resid-obl.png")
ggsave(filename = RF_resid_file, plot = RF_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
RF_pred_resid <- ggplot2::ggplot(RF_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Random Forest Fit: Residuals vs Fitted Obligated Funds",
                                x = "Fitted Requested Funds ($)",
                                y = "Residual")
RF_pred_resid

#save file
RF_pred_resid_file = here("results","RF-pred-resid-obl.png")
ggsave(filename = RF_pred_resid_file, plot = RF_pred_resid)

#print the 10 best performing hyperparameter sets
RF_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err)

#print the best model
RF_performance <- RF_res %>% tune::show_best(n = 1)
RF_performance

#compare model performance to null model (and other models)
RF_RMSE <- RF_res %>%
              tune::show_best(n = 1) %>%
              dplyr::transmute(
                rmse = round(mean, 3),
                SE = round(std_err, 4),
                model = "RF") %>%
             dplyr::bind_rows(tree_RMSE)
RF_RMSE
```

---

## Bagged Decision Tree Model

<br>

### 1. Model Specification
```{r}
#run parallels to determine number of cores
cores <- parallel::detectCores() - 1
cores

cl <- makeCluster(cores)

registerDoParallel(cl)

#start over with data split to allow for bootstrapping
#set random seed
set.seed(20140102)

#use bootstrap instead of CV
obl_data_bootsplit <- obl_train_data %>%
                        rsample::bootstraps(times = 10,
                                            strata = "incidentType")

#define model
bag_tree_mod <- baguette::bag_tree(cost_complexity = tune(),
                                   tree_depth = tune(),
                                   min_n = tune()) %>%
                parsnip::set_engine("rpart",
                                    times = 10) %>%
                parsnip::set_mode("regression")

#create recipe that codes categorical variables as dummy variables
BT_rec <- recipes::recipe(OblAmt ~ ., data = obl_train_data) %>%
          step_string2factor(all_nominal())
```
---

### 2. Workflow Definition
```{r}
#define workflow for tree
BT_wflow <- workflows::workflow() %>%
               workflows::add_model(bag_tree_mod) %>%
               workflows::add_recipe(BT_rec)
```

---

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
tree_grid <- dials::grid_regular(cost_complexity(),
                                 tree_depth(),
                                 min_n(),
                                 levels = 5)
```

---

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
BT_res <- baguette::bag_tree(cost_complexity = tune(),
                                   tree_depth = tune(),
                                   min_n = tune()) %>%
          parsnip::set_engine("rpart",
                              times = 10) %>%
          parsnip::set_mode("regression") %>%
          tune::tune_grid(preprocessor = BT_rec,
                          resamples = obl_data_bootsplit,
                          grid = tree_grid,
                          metrics = metric_set(rmse))

#collect metrics
BT_res %>% workflowsets::collect_metrics()

#default visualization
BT_auto_plot <- BT_res %>% autoplot()
BT_auto_plot

#save file
BT_auto_fig_file = here("results","BT-auto-plot-obl.png")
ggsave(filename = BT_auto_fig_file, plot = BT_auto_plot)

#more detailed plot
BT_det_plot <- BT_res %>%
                    workflowsets::collect_metrics() %>%
                    dplyr::mutate(tree_depth = factor(tree_depth)) %>%
                    ggplot2::ggplot(aes(cost_complexity, mean, color = tree_depth)) +
                             geom_line(size = 1.5, alpha = 0.6) +
                             geom_point(size = 2) +
                             facet_wrap(~ .metric, scales = "free", nrow = 2) +
                             scale_x_log10(labels = scales::label_number()) +
                             scale_color_viridis_d(option = "plasma", begin = 0.9, end = 0)
BT_det_plot

#save file
BT_det_fig_file = here("results","BT-det-plot-obl.png")
ggsave(filename = BT_det_fig_file, plot = BT_det_plot)
```

---

### 5. Identify Best Model
```{r}
#select the BT model with the lowest rmse
BT_lowest_rmse <- BT_res %>%
                        tune::select_best("rmse")

#finalize the workflow by using the selected BT model
best_BT_wflow <- BT_wflow %>%
                      tune::finalize_workflow(BT_lowest_rmse)
best_BT_wflow

#one last fit on the training data
best_BT_fit <- best_BT_wflow %>%
                    parsnip::fit(data = obl_train_data)
```

---

### 6. Model evaluation
```{r}
#extract model from final fit
x_BT <- best_BT_fit$fit$fit$fit

#vip package doesn't support BT

#find predictions and intervals
BT_resid <- best_BT_fit %>%
                  broom.mixed::augment(new_data = obl_train_data) %>%
                  dplyr::select(.pred, OblAmt) %>%
                  dplyr::mutate(.resid = .pred - OblAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
BT_pred_act <- ggplot2::ggplot(BT_resid, aes(x = .pred, y = OblAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point() +
                        labs(title = "Bagged Tree Fit: Predicted vs. Actual Obligated Funds",
                              x = "Predicted Obligated Funds ($)",
                              y = "Actual Obligated Funds ($)")
BT_pred_act

#save file
BT_pred_act_file = here("results","BT-pred-act-obl.png")
ggsave(filename = BT_pred_act_file, plot = BT_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
BT_resids <- ggplot2::ggplot(BT_resid, aes(x = as.numeric(row.names(BT_resid)), y = .resid))+
                      geom_hline(yintercept = 0, color = "red", lty = 2) +
                      geom_point() +
                      labs(title = "Bagged Tree Fit: Residuals",
                            x = "Observation Number",
                            y = "Residual")
BT_resids

#save file
BT_resid_file = here("results","BT-resids-obl.png")
ggsave(filename = BT_resid_file, plot = BT_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
BT_pred_resid <- ggplot2::ggplot(BT_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Bagged Tree Fit: Residuals vs Fitted Obligated Funds",
                                x = "Predicted Obligated Funds ($)",
                                y = "Residual")
BT_pred_resid

#save file
BT_pred_resid_file = here("results","BT-pred-resid-obl.png")
ggsave(filename = BT_pred_resid_file, plot = BT_pred_resid)

#print model performance
#print 10 best performing hyperparameter sets
BT_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err, cost_complexity) %>%
  dplyr::mutate(rmse = round(rmse, 3),
                std_err = round(std_err, 4),
                cost_complexity = scales::scientific(cost_complexity))

#print the best model performance
BT_performance <- BT_res %>% tune::show_best(n = 1)
print(BT_performance)

#compare model performance to null model
BT_RMSE <- BT_res %>%
                tune::show_best(n = 1) %>%
                dplyr::transmute(
                  rmse = round(mean, 3),
                  SE = round(std_err, 4),
                  model = "Bagged Tree") %>%
               dplyr::bind_rows(RF_RMSE)
BT_RMSE

#save combined table
model_comp_file = here("results", "ML-model-comp-obl.Rds")
saveRDS(BT_RMSE, file = model_comp_file)
```

In examining the models, the Bagged Tree has the best RMSE, but it also has a higher SE than the other models. We will still choose the bagged tree as our model on which to evaluate the test data.

--

## Model Selection and Evaluation
Using the bagged tree model
```{r}
#fit model to training set but evaluate on the test data
BT_fit_test <- best_BT_wflow %>%
                    tune::last_fit(split = obl_data_split)

#compare test performance against training performance
BT_rmse_test <- collect_metrics(BT_fit_test) %>%
  dplyr::select(rmse = .estimate) %>%
  dplyr::mutate(data = "test")

BT_comp <- BT_RMSE %>%
                  dplyr::filter(model == "Bagged Tree") %>%
                  dplyr::transmute(rmse, data = "train") %>%
                  dplyr::bind_rows(BT_rmse_test) %>%
                  dplyr::slice(-3) #don't know why the third row shows up
BT_comp
#RMSEs are incredibly different...

#find predictions and intervals
BT_resid_fit <- BT_fit_test %>%
                  broom.mixed::augment() %>%
                  dplyr::select(.pred, OblAmt) %>%
                  dplyr::mutate(.resid = .pred - OblAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
BT_pred_act_test <- ggplot2::ggplot(BT_resid_fit, aes(x = .pred, y = OblAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point() +
                        labs(title = "Bagged Tree Test Fit: Predicted vs. Actual Obligated Funds",
                              x = "Predicted Obligated Funds ($)",
                              y = "Actual Obligated Funds ($)")
BT_pred_act_test

#save file
BT_pred_act_test_file = here("results","BT-pred-act-test-obl.png")
ggsave(filename = BT_pred_act_test_file, plot = BT_pred_act_test)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
BT_resids_test <- ggplot2::ggplot(BT_resid_fit, aes(x = as.numeric(row.names(BT_resid_fit)), y = .resid))+
                      geom_hline(yintercept = 0, color = "red", lty = 2) +
                      geom_point() +
                      labs(title = "Bagged Tree Test Fit: Residuals",
                            x = "Observation Number",
                            y = "Residual")
BT_resids_test

#save file
BT_resid_test_file = here("results","BT-resid-test-obl.png")
ggsave(filename = BT_resid_test_file, plot = BT_resids_test)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
BT_pred_resid_test <- ggplot2::ggplot(BT_resid_fit, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Bagged Tree Test Fit: Residuals vs Fitted Obligated Funds",
                                x = "Predicted Obligated Funds ($)",
                                y = "Residual")
BT_pred_resid_test

#save file
BT_pred_resid_test_file = here("results","BT-pred-resid-test-obl.png")
ggsave(filename = BT_pred_resid_test_file, plot = BT_pred_resid_test)
```