### 2021-11-28 code/analysis_code/analysis-ML-RA-outlier.Rmd
Outlier Analysis

---
title: "Analysis: Machine Learning - ReqAmt"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---

---

## Introduction

This script uses requested FEMA funds as the outcome of interest and fits the following models to the analysis data:

* Null
* Decision Tree
* Random Forest
* Bagged Tree

It compares the  models, and then finally fits the “best” model to the test data.

<br>

Follow the previous processing and analysis markdowns to generate the analysis data used here.

<br>

Each model will follow this process:

1. Model Specification
2. Workflow Definition
3. Tuning Grid Specification
4. Tuning Using Cross-Validation and the tune_grid() function
5. Identify Best Model
6. Model Evaluation

---

## Required Packages
The following R packages are required for this script:

* here: for data loading/saving
* tidyverse: for data management
* tidymodels: for data modeling
* skimr: for variable summaries
* broom.mixed: for converting bayesian models to tidy tibbles
* rpart.plot: for visualizing a decision tree
* vip: for variable importance plots
* glmnet: for lasso models
* doParallel: for parallel backend for tuning processes
* ranger: for random forest models
* baguette: for bagged decision tree
```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(rpart.plot) #for visualizing a decision tree
library(vip) #for variable importance plots
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(baguette) #for bagged decision tree

#global environment options
# formatting for script to avoid scientific notation output
options(scipen=999)
```

---

## Load Data
Load the analysis data from the `processed_data` folder in the project file.
```{r load data}
#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","analysisdata.rds")

#load data. 
analysisdata <- readRDS(data_location)

#summary of data using skimr package
skimr::skim(analysisdata)
```

---

## Data Setup
We will specify the following parameters:

* Drop the unused variables
* Perform feature engineering
* Transform Incident Month into ordinal variable
* Remove outlier
* Set the random seed to 123
* Split the dataset into 70% training, 30% testing, stratified on incidentType
* 5-fold cross validation, 5 times repeated, stratified on incidentType for the CV folds
* Create a recipe for data and fitting that codes categorical variables as dummy variables
```{r data setup}
#requested amount filter
req_data <- analysisdata %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, ResponseDays, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, ReqAmt)

#transform incident month into ordinal variable
req_data$IncidentMonth <- ordered(req_data$IncidentMonth, levels = 1:12,
                                  labels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))

#not enough values per state, so create a variable that is FEMA Regions
req_data$FEMARegion <- ifelse((req_data$state == "CT" |
                                 req_data$state == "ME" |
                                 req_data$state == "MA" |
                                 req_data$state == "NH" |
                                 req_data$state == "RI" |
                                 req_data$state == "VT"),
                              "Region I",
                              ifelse((req_data$state == "NJ" |
                                         req_data$state == "NY" |
                                         req_data$state == "PR" |
                                         req_data$state == "VI"),
                                      "Region II",
                                       ifelse((req_data$state == "DE" |
                                                 req_data$state == "MD" |
                                                 req_data$state == "PA" |
                                                 req_data$state == "VA" |
                                                 req_data$state == "DC" |
                                                 req_data$state == "WV"),
                                              "Region III",
                                               ifelse((req_data$state == "AL" |
                                                         req_data$state == "FL" |
                                                         req_data$state == "GA" |
                                                         req_data$state == "KY" |
                                                         req_data$state == "MS" |
                                                         req_data$state == "NC" |
                                                         req_data$state == "SC" |
                                                         req_data$state == "TN"),
                                                      "Region IV",
                                                      ifelse((req_data$state == "IL" |
                                                               req_data$state == "IN" |
                                                               req_data$state == "MI" |
                                                               req_data$state == "MN" |
                                                               req_data$state == "OH" |
                                                               req_data$state == "WI"),
                                                            "Region V",
                                                             ifelse((req_data$state == "AR" |
                                                                         req_data$state == "LA" |
                                                                         req_data$state == "NM" |
                                                                         req_data$state == "OK" |
                                                                         req_data$state == "TX"),
                                                                      "Region VI",
                                                                       ifelse((req_data$state == "IA" |
                                                                                 req_data$state == "KS" |
                                                                                 req_data$state == "MO" |
                                                                                 req_data$state == "NE"),
                                                                              "Region VII",
                                                                               ifelse((req_data$state == "CO" |
                                                                                         req_data$state == "MT" |
                                                                                         req_data$state == "ND" |
                                                                                         req_data$state == "SD" |
                                                                                         req_data$state == "UT" |
                                                                                         req_data$state == "WY"),
                                                                                      "Region VIII",
                                                                                       ifelse((req_data$state == "AZ" |
                                                                                                 req_data$state == "CA" |
                                                                                                 req_data$state == "HI" |
                                                                                                 req_data$state == "NV"),
                                                                                              "Region IX",
                                                                                               ifelse((req_data$state == "AK" |
                                                                                                         req_data$state == "ID" |
                                                                                                         req_data$state == "OR" |
                                                                                                         req_data$state == "WA"),
                                                                                                      "Region X",
                                                                                                      NA
                                                                                                      )
                                                                                              )
                                                                                      )
                                                                              )
                                                                      )
                                                            )
                                                      )
                                              )
                                     )
                              )

#drop state variable
req_data <- req_data %>%
              dplyr::select(-c(state))

#add to "other" category for incident Types that are infrequent
req_data$incidentType[(req_data$incidentType == "Chemical" |
                        req_data$incidentType == "Coastal Storm" |
                        req_data$incidentType == "Dam/Levee Break" |
                        req_data$incidentType == "Mud/Landslide" |
                        req_data$incidentType == "Volcano" |
                        req_data$incidentType == "Terrorist" |
                        req_data$incidentType == "Snow")] <- "Other"

#remove Hurricane Maria outlier
req_data <-subset(req_data, ReqAmt< 4000000000)

#set random seed
set.seed(123)

#split dataset into 66.6% training, 33.3% testing
#use incidentType as stratification
req_data_split <- rsample::initial_split(req_data, prop = 2/3,
                                     strata = incidentType)

#create dataframes for the two sets:
req_train_data <- rsample::training(req_data_split)
req_test_data <- rsample::testing(req_data_split)

#5-fold cross validation, repeated 5 times, stratified by incidentType
req_folds <- rsample::vfold_cv(req_train_data,
                               strata = incidentType,
                               v = 5,
                               repeats = 5)

#create recipe that codes categorical variables as dummy variables
RA_rec <- recipes::recipe(ReqAmt ~ ., data = req_train_data) %>%
          recipes::step_dummy(all_nominal_predictors())
```

---

## Null Model
Determine the performance of a null model (i.e. one with no predictors) and compute the RMSE for both training and test data.
```{r null model}
#create null model
null_mod <- parsnip::null_model() %>%
            parsnip::set_engine("parsnip") %>%
            parsnip::set_mode("regression")

#add recipe and model into workflow
null_wflow <- workflows::workflow() %>%
              workflows::add_recipe(RA_rec) %>%
              workflows::add_model(null_mod)

#"fit" model to training data
req_null_train <- null_wflow %>%
                  parsnip::fit(data = req_train_data)

#summary of null model with training data to get mean (which in this case is the RMSE)
req_null_train_sum <- broom.mixed::tidy(req_null_train)
req_null_train_sum

#RMSE for training data for formatting later
req_null_RMSE_train <- tibble::tibble(
                          rmse = rmse_vec(truth = req_train_data$ReqAmt,
                                              estimate = rep(mean(req_train_data$ReqAmt), nrow(req_train_data))),
                          SE = 0,
                          model = "Null - Train")

#"fit" model to test data
req_null_test <- null_wflow %>%
                  parsnip::fit(data = req_test_data)

#summary of null model with test data to get mean (which in this case is the RMSE)
req_null_test_sum <- broom.mixed::tidy(req_null_test)
req_null_test_sum

#RMSE for testing data for formatting later
req_null_RMSE_test <- tibble::tibble(
                          rmse = rmse_vec(truth = req_test_data$ReqAmt,
                                              estimate = rep(mean(req_test_data$ReqAmt), nrow(req_test_data))),
                          SE = 0,
                          model = "Null - Test")
```

---

## Decision Tree

<br>

### 1. Model Specification
```{r}
#run parallels to determine number of cores
cores <- parallel::detectCores() - 1
cores

cl <- makeCluster(cores)

registerDoParallel(cl)

#define the tree model
tree_mod <-
  parsnip::decision_tree(cost_complexity = tune(),
                         tree_depth = tune(),
                         min_n = tune()) %>%
  parsnip::set_engine("rpart") %>%
  parsnip::set_mode("regression")

#use the recipe specified earlier (line 133)
```
---

### 2. Workflow Definition
```{r}
#define workflow for tree
tree_wflow <- workflows::workflow() %>%
               workflows::add_model(tree_mod) %>%
               workflows::add_recipe(RA_rec)
```

---

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
tree_grid <- dials::grid_regular(cost_complexity(),
                                 tree_depth(),
                                 min_n(),
                                 levels = 5)

#tree depth
tree_grid %>%
  dplyr::count(tree_depth)
```

---

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
tree_res <- tree_wflow %>%
                tune::tune_grid(resamples = req_folds,
                                grid = tree_grid,
                                control = control_grid(verbose = TRUE),
                                metrics = yardstick::metric_set(rmse))

#collect metrics
tree_res %>% workflowsets::collect_metrics()

#default visualization
DT_auto_plot <- tree_res %>% autoplot()
DT_auto_plot

#save file
DT_auto_fig_file = here("results","DT-auto-plot.png")
ggsave(filename = DT_auto_fig_file, plot = DT_auto_plot)

#more detailed plot
DT_detail_plot <- tree_res %>%
                    workflowsets::collect_metrics() %>%
                    dplyr::mutate(tree_depth = factor(tree_depth)) %>%
                    ggplot2::ggplot(aes(cost_complexity, mean, color = tree_depth)) +
                             geom_line(size = 1.5, alpha = 0.6) +
                             geom_point(size = 2) +
                             facet_wrap(~ .metric, scales = "free", nrow = 2) +
                             scale_x_log10(labels = scales::label_number()) +
                             scale_color_viridis_d(option = "plasma", begin = 0.9, end = 0)
DT_detail_plot

#save file
DT_det_fig_file = here("results","DT-det-plot.png")
ggsave(filename = DT_det_fig_file, plot = DT_detail_plot)

```

---

### 5. Identify Best Model
```{r}
#select the tree model with the lowest rmse
tree_lowest_rmse <- tree_res %>%
                        tune::select_best("rmse")

#finalize the workflow by using the selected DT model
best_tree_wflow <- tree_wflow %>%
                      tune::finalize_workflow(tree_lowest_rmse)
best_tree_wflow

#one last fit on the training data
best_tree_fit <- best_tree_wflow %>%
                    parsnip::fit(data = req_train_data)
```

---

### 6. Model evaluation
```{r}
#plot the tree
rpart.plot::rpart.plot(x = workflowsets::extract_fit_parsnip(best_tree_fit)$fit,
                                                     roundint = F,
                                                     type = 5,
                                                     digits = 5,
                                                     main = "Selected Decision Tree Model")
#save manually -- ggsave throws an error

#find predictions and intervals
tree_resid <- best_tree_fit %>%
                  broom.mixed::augment(new_data = req_train_data) %>%
                  dplyr::select(.pred, ReqAmt) %>%
                  dplyr::mutate(.resid = .pred - ReqAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
DT_pred_act <- ggplot2::ggplot(tree_resid, aes(x = .pred, y = ReqAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point() +
                        labs(title = "Decision Tree Fit: Predicted vs. Actual Requested Funds",
                              x = "Predicted Requested Funds ($)",
                              y = "Actual Requested Funds ($)")
DT_pred_act

#save file
DT_pred_act_file = here("results","DT-pred-act.png")
ggsave(filename = DT_pred_act_file, plot = DT_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
DT_resid <- ggplot2::ggplot(tree_resid, aes(x = as.numeric(row.names(tree_resid)), y = .resid))+
                     geom_hline(yintercept = 0, color = "red", lty = 2) +
                     geom_point() +
                     labs(title = "Decision Tree Fit: Residuals",
                          x = "Observation Number",
                          y = "Residual")
DT_resid

#save file
DT_resid_file = here("results","DT-resid.png")
ggsave(filename = DT_resid_file, plot = DT_resid)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
DT_pred_resid <- ggplot2::ggplot(tree_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Decision Tree Fit: Residuals vs Fitted Requested Funds",
                                x = "Predicted Requested Funds ($)",
                                y = "Residual")
DT_pred_resid

#save file
DT_pred_resid_file = here("results","DT-pred-resid.png")
ggsave(filename = DT_pred_resid_file, plot = DT_pred_resid)

#print model performance
#print 10 best performing hyperparameter sets
tree_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err, cost_complexity) %>%
  dplyr::mutate(rmse = round(rmse, 3),
                std_err = round(std_err, 4),
                cost_complexity = scales::scientific(cost_complexity))

#print the best model performance
tree_performance <- tree_res %>% tune::show_best(n = 1)
print(tree_performance)

#compare model performance to null model
tree_RMSE <- tree_res %>%
                tune::show_best(n = 1) %>%
                dplyr::transmute(
                  rmse = round(mean, 3),
                  SE = round(std_err, 4),
                  model = "Tree") %>%
               dplyr::bind_rows(req_null_RMSE_train)
tree_RMSE
```

The identified best performing tree model has some interesting results. It identifies incident duration, state population, total number of federal agencies, total number of state counties, and individual/households program awardance as significant decision points for predicting requesting FEMA funding. All of these variables are to be expected, but it is interesting that incident type or declaration type are not included as significant. The RMSE is improved from the Null model, but there's still opportunity for improvement. Let's try another model.

<br>

---

## Random Forest Model

<br>

### 1. Model Specification
```{r}
#run parallels to determine number of cores
cores <- parallel::detectCores() - 1
cores

cl <- makeCluster(cores)

registerDoParallel(cl)

#define the RF model
RF_mod <-
  parsnip::rand_forest(mtry = tune(),
                       min_n = tune(),
                       trees = tune()) %>%
  parsnip::set_engine("ranger",
                      importance = "permutation") %>%
  parsnip::set_mode("regression")

#use the recipe specified earlier (line 133)

#check to make sure identified parameters will be tuned
RF_mod %>% tune::parameters()
```

---

### 2. Workflow Definition
```{r}
#define workflow for RF regression
RF_wflow <- workflows::workflow() %>%
               workflows::add_model(RF_mod) %>%
               workflows::add_recipe(RA_rec)
```

---

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
RF_grid <- expand.grid(mtry = c(3, 4, 5, 6),
                       min_n = c(40, 50, 60),
                       trees = c(500,1000))
```

---

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
RF_res <- RF_wflow %>%
              tune::tune_grid(resamples = req_folds,
                              grid = RF_grid,
                              control = control_grid(verbose = TRUE, save_pred = TRUE),
                              metrics = metric_set(rmse))

#look at top 5 RF models
top_RF_models <- RF_res %>%
                    tune::show_best("rmse", n = 5)
top_RF_models

#default visualization
RF_auto_plot <- RF_res %>% autoplot()
RF_auto_plot

#save file
RF_auto_fig_file = here("results","RF-auto-plot.png")
ggsave(filename = RF_auto_fig_file, plot = RF_auto_plot)
```

---

### 5. Identify Best Model
```{r}
#select the RF model with the lowest rmse
RF_lowest_rmse <- RF_res %>%
                      tune::select_best("rmse")

#finalize the workflow by using the selected RF model
best_RF_wflow <- RF_wflow %>%
                      tune::finalize_workflow(RF_lowest_rmse)
best_RF_wflow

#one last fit on the training data
best_RF_fit <- best_RF_wflow %>%
                    parsnip::fit(data = req_train_data)
```

---

### 6. Model evaluation
```{r}
#extract model from final fit
x_RF <- best_RF_fit$fit$fit$fit

#plot most important predictors in the model
RF_plot <- vip::vip(x_RF, num_features = 20)
RF_plot

#save manually-- ggsave throws errors

#find predictions and intervals
RF_resid <- best_RF_fit %>%
                broom.mixed::augment(new_data = req_train_data) %>%
                dplyr::select(.pred, ReqAmt) %>%
                dplyr::mutate(.resid = .pred - ReqAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline is a 45 degree line, along which the results should fall
RF_pred_act <- ggplot2::ggplot(RF_resid, aes(x = .pred, y = ReqAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) +
                        geom_point() +
                        labs(title = "RF Fit: Actual vs. Predicted Requested Funds",
                              x = "Predicted Requested Funds ($)",
                              y = "Actual Requested Funds ($)")
RF_pred_act

#save file
RF_pred_act_file = here("results","RF-pred-act.png")
ggsave(filename = RF_pred_act_file, plot = RF_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
RF_resids <- ggplot2::ggplot(RF_resid, aes(x = as.numeric(row.names(RF_resid)), y = .resid))+
                            geom_hline(yintercept = 0, color = "red", lty = 2) +
                            geom_point() +
                            labs(title = "RF Fit: Residuals",
                                  x = "Observation Number",
                                  y = "Residual")
RF_resids

#save file
RF_resid_file = here("results","RF-resid.png")
ggsave(filename = RF_resid_file, plot = RF_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
RF_pred_resid <- ggplot2::ggplot(RF_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Random Forest Fit: Residuals vs Fitted Requested Funds",
                                x = "Fitted Requested Funds ($)",
                                y = "Residual")
RF_pred_resid

#save file
RF_pred_resid_file = here("results","RF-pred-resid.png")
ggsave(filename = RF_pred_resid_file, plot = RF_pred_resid)

#print the 10 best performing hyperparameter sets
RF_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err)

#print the best model
RF_performance <- RF_res %>% tune::show_best(n = 1)
RF_performance

#compare model performance to null model (and other models)
RF_RMSE <- RF_res %>%
              tune::show_best(n = 1) %>%
              dplyr::transmute(
                rmse = round(mean, 3),
                SE = round(std_err, 4),
                model = "RF") %>%
             dplyr::bind_rows(tree_RMSE)
RF_RMSE
```

In examining the results of the RF model within the context of RMSE, it performs marginally better than the decision tree. Interestingly, it includes declaration month of Jan - March as more important than state population. Let's try one more model.

---

## Bagged Decision Tree Model

<br>

### 1. Model Specification
```{r}
#run parallels to determine number of cores
cores <- parallel::detectCores() - 1
cores

cl <- makeCluster(cores)

registerDoParallel(cl)

#start over with data split to allow for bootstrapping
#set random seed
set.seed(20140102)

#use bootstrap instead of CV
req_data_bootsplit <- req_train_data %>%
                        rsample::bootstraps(times = 10,
                                            strata = "incidentType")

#define model
bag_tree_mod <- baguette::bag_tree(cost_complexity = tune(),
                                   tree_depth = tune(),
                                   min_n = tune()) %>%
                parsnip::set_engine("rpart",
                                    times = 10) %>%
                parsnip::set_mode("regression")

#create recipe that codes categorical variables as dummy variables
BT_rec <- recipes::recipe(ReqAmt ~ ., data = req_train_data) %>%
          step_string2factor(all_nominal())
```
---

### 2. Workflow Definition
```{r}
#define workflow for tree
BT_wflow <- workflows::workflow() %>%
               workflows::add_model(bag_tree_mod) %>%
               workflows::add_recipe(BT_rec)
```

---

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
tree_grid <- dials::grid_regular(cost_complexity(),
                                 tree_depth(),
                                 min_n(),
                                 levels = 5)
```

---

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
BT_res <- BT_wflow %>%
          tune::tune_grid(resamples = req_data_bootsplit,
                          grid = tree_grid,
                          metrics = metric_set(rmse),
                          control = control_resamples(save_pred = TRUE,
                                                      extract = function(x) extract_model(x)))

#collect metrics
BT_res %>% workflowsets::collect_metrics()

#default visualization
BT_auto_plot <- BT_res %>% autoplot()
BT_auto_plot

#save file
BT_auto_fig_file = here("results","BT-auto-plot.png")
ggsave(filename = BT_auto_fig_file, plot = BT_auto_plot)

#more detailed plot
BT_det_plot <- BT_res %>%
                    workflowsets::collect_metrics() %>%
                    dplyr::mutate(tree_depth = factor(tree_depth)) %>%
                    ggplot2::ggplot(aes(cost_complexity, mean, color = tree_depth)) +
                             geom_line(size = 1.5, alpha = 0.6) +
                             geom_point(size = 2) +
                             facet_wrap(~ .metric, scales = "free", nrow = 2) +
                             scale_x_log10(labels = scales::label_number()) +
                             scale_color_viridis_d(option = "plasma", begin = 0.9, end = 0)
BT_det_plot

#save file
BT_det_fig_file = here("results","BT-det-plot.png")
ggsave(filename = BT_det_fig_file, plot = BT_det_plot)
```

---

### 5. Identify Best Model
```{r}
#select the BT model with the lowest rmse
BT_lowest_rmse <- BT_res %>%
                        tune::select_best("rmse")

#finalize the workflow by using the selected BT model
best_BT_wflow <- BT_wflow %>%
                      tune::finalize_workflow(BT_lowest_rmse)
best_BT_wflow

#one last fit on the training data
best_BT_fit <- best_BT_wflow %>%
                    parsnip::fit(data = req_train_data)
```

---

### 6. Model evaluation
```{r}
#bag roots function to extract most important variables
bag_roots <-  function(x){
  x %>% 
    dplyr::select(.extracts) %>% 
    unnest(cols = c(.extracts)) %>% 
    dplyr::mutate(models = map(.extracts,
                               ~.x$model_df)) %>% 
    dplyr::select(-.extracts) %>% 
    unnest(cols = c(models)) %>% 
    mutate(root = map_chr(model,
                          ~as.character(.x$fit$frame[1, 1]))) %>%
    dplyr::select(root)  
}

#plot most important variables
bag_roots(BT_res) %>% 
  ggplot(mapping = aes(x = fct_rev(fct_infreq(root)))) + 
  geom_bar() + 
  coord_flip() + 
  labs(x = "root", y = "count")

#find predictions and intervals
BT_resid <- best_BT_fit %>%
                  broom.mixed::augment(new_data = req_train_data) %>%
                  dplyr::select(.pred, ReqAmt) %>%
                  dplyr::mutate(.resid = .pred - ReqAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
BT_pred_act <- ggplot2::ggplot(BT_resid, aes(x = .pred, y = ReqAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point() +
                        labs(title = "Bagged Tree Fit: Predicted vs. Actual Requested Funds",
                              x = "Predicted Requested Funds ($)",
                              y = "Actual Requested Funds ($)")
BT_pred_act

#save file
BT_pred_act_file = here("results","BT-pred-act.png")
ggsave(filename = BT_pred_act_file, plot = BT_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
BT_resids <- ggplot2::ggplot(BT_resid, aes(x = as.numeric(row.names(BT_resid)), y = .resid))+
                      geom_hline(yintercept = 0, color = "red", lty = 2) +
                      geom_point() +
                      labs(title = "Bagged Tree Fit: Residuals",
                            x = "Observation Number",
                            y = "Residual")
BT_resids

#save file
BT_resid_file = here("results","BT-resid.png")
ggsave(filename = BT_resid_file, plot = BT_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
BT_pred_resid <- ggplot2::ggplot(BT_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Bagged Tree Fit: Residuals vs Fitted Requested Funds",
                                x = "Predicted Requested Funds ($)",
                                y = "Residual")
BT_pred_resid

#save file
BT_pred_resid_file = here("results","BT-pred-resid.png")
ggsave(filename = BT_pred_resid_file, plot = BT_pred_resid)

#print model performance
#print 10 best performing hyperparameter sets
BT_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err, cost_complexity) %>%
  dplyr::mutate(rmse = round(rmse, 3),
                std_err = round(std_err, 4),
                cost_complexity = scales::scientific(cost_complexity))

#print the best model performance
BT_performance <- BT_res %>% tune::show_best(n = 1)
print(BT_performance)

#compare model performance to null model
BT_RMSE <- BT_res %>%
                tune::show_best(n = 1) %>%
                dplyr::transmute(
                  rmse = round(mean, 3),
                  SE = round(std_err, 4),
                  model = "Bagged Tree") %>%
               dplyr::bind_rows(RF_RMSE)
BT_RMSE

#save combined table
model_comp_file = here("results", "ML-model-comp.Rds")
saveRDS(BT_RMSE, file = model_comp_file)
```

In examining the models, the Bagged Tree has the best RMSE, but it also has a higher SE than the other models. We will still choose the bagged tree as our model on which to evaluate the test data.

--

## Model Selection and Evaluation
Using the bagged tree model
```{r}
#fit model to training set but evaluate on the test data
BT_fit_test <- best_BT_wflow %>%
                    tune::last_fit(split = req_data_split)

#compare test performance against training performance
BT_rmse_test <- collect_metrics(BT_fit_test) %>%
  dplyr::select(rmse = .estimate) %>%
  dplyr::mutate(data = "test")

BT_comp <- BT_RMSE %>%
                  dplyr::filter(model == "Bagged Tree") %>%
                  dplyr::transmute(rmse, data = "train") %>%
                  dplyr::bind_rows(BT_rmse_test) %>%
                  dplyr::slice(-3) #don't know why the third row shows up
BT_comp
#RMSEs are incredibly different...

#find predictions and intervals
BT_resid_fit <- BT_fit_test %>%
                  broom.mixed::augment() %>%
                  dplyr::select(.pred, ReqAmt) %>%
                  dplyr::mutate(.resid = .pred - ReqAmt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
BT_pred_act_test <- ggplot2::ggplot(BT_resid_fit, aes(x = .pred, y = ReqAmt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point() +
                        labs(title = "Bagged Tree Test Fit: Predicted vs. Actual Requested Funds",
                              x = "Predicted Requested Funds ($)",
                              y = "Actual Requested Funds ($)")
BT_pred_act_test

#save file
BT_pred_act_test_file = here("results","BT-pred-act-test.png")
ggsave(filename = BT_pred_act_test_file, plot = BT_pred_act_test)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
BT_resids_test <- ggplot2::ggplot(BT_resid_fit, aes(x = as.numeric(row.names(BT_resid_fit)), y = .resid))+
                      geom_hline(yintercept = 0, color = "red", lty = 2) +
                      geom_point() +
                      labs(title = "Bagged Tree Test Fit: Residuals",
                            x = "Observation Number",
                            y = "Residual")
BT_resids_test

#save file
BT_resid_test_file = here("results","BT-resid-test.png")
ggsave(filename = BT_resid_test_file, plot = BT_resids_test)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
BT_pred_resid_test <- ggplot2::ggplot(BT_resid_fit, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point() +
                          labs(title = "Bagged Tree Test Fit: Residuals vs Fitted Requested Funds",
                                x = "Predicted Requested Funds ($)",
                                y = "Residual")
BT_pred_resid_test

#save file
BT_pred_resid_test_file = here("results","BT-pred-resid-test.png")
ggsave(filename = BT_pred_resid_test_file, plot = BT_pred_resid_test)
```