### 2021-12-12 code/analysis_code/analysis-ML-RA.Rmd

---
title: "Analysis: Machine Learning - ReqAmt"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---

---

## Introduction

This script uses requested FEMA funds as the outcome of interest and fits the following models to the analysis data:

* Null
* Decision Tree
* Random Forest
* Elastic Net
* LASSO

It compares the  models, and then finally fits the “best” model to the test data.

<br>

Follow the previous processing and analysis markdowns to generate the analysis data used here.

<br>

Each model will follow this process:

1. Model Specification
2. Workflow Definition
3. Tuning Grid Specification
4. Tuning Using Cross-Validation and the tune_grid() function
5. Identify Best Model
6. Model Evaluation

---

## Required Packages
The following R packages are required for this script:

* here: for data loading/saving
* tidyverse: for data management
* tidymodels: for data modeling
* skimr: for variable summaries
* spatialsample: for spatial cross-validation
* broom.mixed: for converting bayesian models to tidy tibbles
* rpart.plot: for visualizing a decision tree
* vip: for variable importance plots
* glmnet: for lasso models
* doParallel: for parallel backend for tuning processes
* ranger: for random forest models
* baguette: for bagged decision tree
* viridis: for a consistent color scheme in figures
```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(spatialsample) #for spatial cross validation
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(rpart.plot) #for visualizing a decision tree
library(vip) #for variable importance plots
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(baguette) #for bagged decision tree
library(viridis) #for a consistent color scheme in figures

#global environment options
# formatting for script to avoid scientific notation output
options(scipen=999)

#set ggplot theme to classic
ggplot2::theme_set(theme_classic())

#center plot titles
theme_update(plot.title = element_text(hjust = 0.5))
```

---

## Load Data
Load the ReqAmt analysis data from the `processed_data` folder in the project file.
```{r load data}
#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","analysisdata-RA.rds")

#load data. 
req_data <- readRDS(data_location)

#summary of data using skimr package
skimr::skim(req_data)
```

---

## Data Setup
We will specify the following parameters:

* Log transform the outcome
* Drop `ReqAmt`, `disasterNumber`, `declarationTitle`
* Set the random seed to 123
* Split the dataset into 70% training, 30% testing
* 5-fold cross validation for the spatial CV folds
* Create a recipe for data and fitting that codes categorical variables as dummy variables
```{r data setup}
#log transform ReqAmt
req_data$logReqAmt <- log(req_data$ReqAmt)

#drop unnecessary variables
req_data <- req_data %>%
              dplyr::select(-c(ReqAmt, disasterNumber, declarationTitle))

#specify all character variables as factors and clean variable names
req_data <- req_data %>% 
              janitor::clean_names() %>%
              dplyr::mutate_if(is.character, factor)

#set random seed
set.seed(123)

#split dataset into 66.6% training, 33.3% testing
#use incidentType as stratification
req_data_split <- rsample::initial_split(req_data, prop = 2/3)

#create dataframes for the two sets:
req_train_data <- rsample::training(req_data_split)
req_test_data <- rsample::testing(req_data_split)

#5-fold spatial cross validation
req_folds_spatial <- spatialsample::spatial_clustering_cv(req_train_data,
                                                          coords = c(latitude, longitude),
                                                          v = 5)

#create recipe that codes categorical variables as dummy variables
RA_rec <- recipes::recipe(log_req_amt ~ ., data = req_train_data) %>%
          recipes::step_unknown(state, new_level = "unknown state") %>%
          recipes::step_dummy(all_nominal_predictors())
```

---

## Null Model
Determine the performance of a null model (i.e. one with no predictors) and compute the RMSE for both training and test data.
```{r null model}
#create null model
null_mod <- parsnip::null_model() %>%
            parsnip::set_engine("parsnip") %>%
            parsnip::set_mode("regression")

#add recipe and model into workflow
null_wflow <- workflows::workflow() %>%
              workflows::add_recipe(RA_rec) %>%
              workflows::add_model(null_mod)

#"fit" model to training data
req_null_train <- null_wflow %>%
                  parsnip::fit(data = req_train_data)

#summary of null model with training data to get mean (which in this case is the RMSE)
req_null_train_sum <- broom.mixed::tidy(req_null_train)
req_null_train_sum

#RMSE for training data for formatting later
req_null_RMSE_train <- tibble::tibble(
                          rmse = rmse_vec(truth = req_train_data$log_req_amt,
                                              estimate = rep(mean(req_train_data$log_req_amt), nrow(req_train_data))),
                          SE = 0,
                          model = "Null - Train")

#"fit" model to test data
req_null_test <- null_wflow %>%
                  parsnip::fit(data = req_test_data)

#summary of null model with test data to get mean (which in this case is the RMSE)
req_null_test_sum <- broom.mixed::tidy(req_null_test)
req_null_test_sum

#RMSE for testing data for formatting later
req_null_RMSE_test <- tibble::tibble(
                          rmse = rmse_vec(truth = req_test_data$log_req_amt,
                                              estimate = rep(mean(req_test_data$log_req_amt), nrow(req_test_data))),
                          SE = 0,
                          model = "Null - Test")
```

---

## Decision Tree

<br>

### 1. Model Specification
```{r}
#run parallels to determine number of cores
cores <- parallel::detectCores() - 1
cores

cl <- makeCluster(cores)

registerDoParallel(cl)

#define the tree model
tree_mod <-
  parsnip::decision_tree(cost_complexity = tune(),
                         tree_depth = tune(),
                         min_n = tune()) %>%
  parsnip::set_engine("rpart") %>%
  parsnip::set_mode("regression")
```

---

### 2. Workflow Definition
```{r}
#define workflow for tree
tree_wflow <- workflows::workflow() %>%
               workflows::add_model(tree_mod) %>%
               workflows::add_recipe(RA_rec)
```

---

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
tree_grid <- dials::grid_regular(cost_complexity(),
                                 tree_depth(),
                                 min_n(),
                                 levels = 5)

#tree depth
tree_grid %>%
  dplyr::count(tree_depth)
```

---

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
tree_res <- tree_wflow %>%
                tune::tune_grid(resamples = req_folds_spatial,
                                grid = tree_grid,
                                control = control_grid(verbose = TRUE),
                                metrics = yardstick::metric_set(rmse))

#collect metrics
tree_res %>% workflowsets::collect_metrics()

#default visualization
DT_auto_plot <- tree_res %>% autoplot()
DT_auto_plot

#save file
DT_auto_fig_file = here("results","DT-auto-plot.png")
ggsave(filename = DT_auto_fig_file, plot = DT_auto_plot)

#more detailed plot
DT_detail_plot <- tree_res %>%
                    workflowsets::collect_metrics() %>%
                    dplyr::mutate(tree_depth = factor(tree_depth)) %>%
                    ggplot2::ggplot(aes(cost_complexity, mean, color = tree_depth)) +
                             geom_line(size = 1.5, alpha = 0.6) +
                             geom_point(size = 2) +
                             facet_wrap(~ .metric, scales = "free") +
                             scale_x_log10(labels = scales::label_number()) +
                             scale_color_viridis_d(option = "plasma", begin = 0.9, end = 0) +
                             labs(x = "cost complexity",
                                  y = "mean",
                                  color = "Tree Depth")
DT_detail_plot

#save file
DT_det_fig_file = here("results","DT-det-plot.png")
ggsave(filename = DT_det_fig_file, plot = DT_detail_plot)

```

---

### 5. Identify Best Model
```{r}
#select the tree model with the lowest rmse
tree_lowest_rmse <- tree_res %>%
                        tune::select_best("rmse")

#finalize the workflow by using the selected DT model
best_tree_wflow <- tree_wflow %>%
                      tune::finalize_workflow(tree_lowest_rmse)
best_tree_wflow

#one last fit on the training data
best_tree_fit <- best_tree_wflow %>%
                    parsnip::fit(data = req_train_data)
```

---

### 6. Model evaluation
```{r}
#plot the tree
rpart.plot::rpart.plot(x = workflowsets::extract_fit_parsnip(best_tree_fit)$fit,
                                                     roundint = F,
                                                     type = 5,
                                                     digits = 5,
                                                     main = "Selected Decision Tree Model",
                                                     box.palette = viridis::viridis(10, option = "D", begin = 0.9, end = 0),
                                                     shadow.col = "grey65",
                                                     col = "grey99")
#save manually as "DT-plot.png" -- ggsave throws an error

#find predictions and intervals
tree_resid <- best_tree_fit %>%
                  broom.mixed::augment(new_data = req_train_data) %>%
                  dplyr::select(.pred, log_req_amt) %>%
                  dplyr::mutate(.resid = .pred - log_req_amt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
DT_pred_act <- ggplot2::ggplot(tree_resid, aes(x = .pred, y = log_req_amt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point(color = "#3b528b") +
                        scale_y_continuous(expand = c(0,0),
                                           limits = c(-30, 25)) +
                        scale_x_continuous(limits = c(0, 25)) +
                        labs(title = "Decision Tree Fit: Predicted vs. Actual Log(Requested Funds)",
                              x = "Predicted Log(Requested Funds) ($)",
                              y = "Actual Log(Requested Funds) ($)") 
DT_pred_act

#save file
DT_pred_act_file = here("results","DT-pred-act.png")
ggsave(filename = DT_pred_act_file, plot = DT_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
DT_resid <- ggplot2::ggplot(tree_resid, aes(x = as.numeric(row.names(tree_resid)), y = .resid))+
                     geom_hline(yintercept = 0, color = "red", lty = 2) +
                     geom_point(color = "#3b528b") +
                     labs(title = "Decision Tree Fit: Residuals",
                          x = "Observation Number",
                          y = "Residual")
DT_resid

#save file
DT_resid_file = here("results","DT-resid.png")
ggsave(filename = DT_resid_file, plot = DT_resid)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
DT_pred_resid <- ggplot2::ggplot(tree_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point(color = "#3b528b") +
                          labs(title = "Decision Tree Fit: Residuals vs Fitted Log(Requested Funds)",
                                x = "Predicted Log(Requested Funds) ($)",
                                y = "Residual")
DT_pred_resid

#save file
DT_pred_resid_file = here("results","DT-pred-resid.png")
ggsave(filename = DT_pred_resid_file, plot = DT_pred_resid)

#print model performance
#print 10 best performing hyperparameter sets
tree_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err, cost_complexity) %>%
  dplyr::mutate(rmse = round(rmse, 3),
                std_err = round(std_err, 4),
                cost_complexity = scales::scientific(cost_complexity))

#print the best model performance
tree_performance <- tree_res %>% tune::show_best(n = 1)
print(tree_performance)

#compare model performance to null model
tree_RMSE <- tree_res %>%
                tune::show_best(n = 1) %>%
                dplyr::transmute(
                  rmse = round(mean, 3),
                  SE = round(std_err, 4),
                  model = "Tree") %>%
               dplyr::bind_rows(req_null_RMSE_train)
tree_RMSE
```

The identified best performing tree model has some interesting results. It identifies percent of state counties awarded IH program, New York, longitude, total number of agencies, FEMA cost share average, total response days, and an incident month of November for predicting requested FEMA funding. New York and November are particularly surprising key predictors. It is interesting that incident type or declaration type are not included as significant. The RMSE actually increases from the Null, which suggests this is not an appropriate model. Moving on.

<br>

---

## Random Forest Model

<br>

### 1. Model Specification
```{r}
#run parallels to determine number of cores
cores <- parallel::detectCores() - 1
cores

cl <- makeCluster(cores)

registerDoParallel(cl)

#define the RF model
RF_mod <-
  parsnip::rand_forest(mtry = tune(),
                       min_n = tune(),
                       trees = tune()) %>%
  parsnip::set_engine("ranger",
                      importance = "permutation") %>%
  parsnip::set_mode("regression")

#use the recipe specified earlier (line 133)

#check to make sure identified parameters will be tuned
RF_mod %>% tune::parameters()
```

---

### 2. Workflow Definition
```{r}
#define workflow for RF regression
RF_wflow <- workflows::workflow() %>%
               workflows::add_model(RF_mod) %>%
               workflows::add_recipe(RA_rec)
```

---

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
RF_grid <- expand.grid(mtry = c(3, 4, 5, 6),
                       min_n = c(40, 50, 60),
                       trees = c(500,1000))
```

---

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
RF_res <- RF_wflow %>%
              tune::tune_grid(resamples = req_folds_spatial,
                              grid = RF_grid,
                              control = control_grid(verbose = TRUE, save_pred = TRUE),
                              metrics = metric_set(rmse))

#look at top 5 RF models
top_RF_models <- RF_res %>%
                    tune::show_best("rmse", n = 5)
top_RF_models

#default visualization
RF_auto_plot <- RF_res %>% autoplot()
RF_auto_plot

#save file
RF_auto_fig_file = here("results","RF-auto-plot.png")
ggsave(filename = RF_auto_fig_file, plot = RF_auto_plot)
```

---

### 5. Identify Best Model
```{r}
#select the RF model with the lowest rmse
RF_lowest_rmse <- RF_res %>%
                      tune::select_best("rmse")

#finalize the workflow by using the selected RF model
best_RF_wflow <- RF_wflow %>%
                      tune::finalize_workflow(RF_lowest_rmse)
best_RF_wflow

#one last fit on the training data
best_RF_fit <- best_RF_wflow %>%
                    parsnip::fit(data = req_train_data)
```

---

### 6. Model evaluation
```{r}
#extract model from final fit
x_RF <- best_RF_fit$fit$fit$fit

#plot most important predictors in the model
RF_plot <- vip::vip(x_RF, 
                    num_features = 20,
                    aesthetics = list(fill = viridis::viridis(20, option = "D", begin = 0.9, end = 0)))
RF_plot

#save manually as "RF-plot.png" -- ggsave throws errors

#find predictions and intervals
RF_resid <- best_RF_fit %>%
                broom.mixed::augment(new_data = req_train_data) %>%
                dplyr::select(.pred, log_req_amt) %>%
                dplyr::mutate(.resid = .pred - log_req_amt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline is a 45 degree line, along which the results should fall
RF_pred_act <- ggplot2::ggplot(RF_resid, aes(x = .pred, y = log_req_amt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) +
                        geom_point(color = "#3b528b") +
                        scale_y_continuous(expand = c(0,0),
                                           limits = c(-30, 25)) +
                        scale_x_continuous(limits = c(0, 25)) +
                        labs(title = "RF Fit: Actual vs. Predicted Log(Requested Funds)",
                              x = "Predicted Log(Requested Funds) ($)",
                              y = "Actual Log(Requested Funds) ($)")
RF_pred_act

#save file
RF_pred_act_file = here("results","RF-pred-act.png")
ggsave(filename = RF_pred_act_file, plot = RF_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
RF_resids <- ggplot2::ggplot(RF_resid, aes(x = as.numeric(row.names(RF_resid)), y = .resid))+
                            geom_hline(yintercept = 0, color = "red", lty = 2) +
                            geom_point(color = "#3b528b") +
                            labs(title = "RF Fit: Residuals",
                                  x = "Observation Number",
                                  y = "Residual")
RF_resids

#save file
RF_resid_file = here("results","RF-resid.png")
ggsave(filename = RF_resid_file, plot = RF_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
RF_pred_resid <- ggplot2::ggplot(RF_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point(color = "#3b528b") +
                          labs(title = "Random Forest Fit: Residuals vs Fitted Log(Requested Funds)",
                                x = "Fitted Log(Requested Funds) ($)",
                                y = "Residual")
RF_pred_resid

#save file
RF_pred_resid_file = here("results","RF-pred-resid.png")
ggsave(filename = RF_pred_resid_file, plot = RF_pred_resid)

#print the 10 best performing hyperparameter sets
RF_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err)

#print the best model
RF_performance <- RF_res %>% tune::show_best(n = 1)
RF_performance

#compare model performance to null model (and other models)
RF_RMSE <- RF_res %>%
              tune::show_best(n = 1) %>%
              dplyr::transmute(
                rmse = round(mean, 3),
                SE = round(std_err, 4),
                model = "RF") %>%
             dplyr::bind_rows(tree_RMSE)
RF_RMSE
```

In examining the results of the RF model within the context of RMSE, it performs marginally better than the decision tree. It identifies total number of federal agencies involved, percent of state counties awarded IH, FEMA cost share average, percent of counties awarded HM, and whether or not IH was awarded to any part of the state as the five most important predictors. Let's try another model to see if we can find an improvement in the RMSE.

---

## Elastic Net Model

<br>

### 1. Model Specification
```{r}
#run parallels to determine number of cores
cores <- parallel::detectCores() - 1
cores

cl <- makeCluster(cores)

registerDoParallel(cl)

#define the elastic net model
elastic_mod <- linear_reg(mode = "regression",
                           penalty = tune(),
                           mixture = tune()) %>%
                parsnip::set_engine("glmnet")
```

---

### 2. Workflow Definition
```{r}
#define elastic net workflow
elastic_wflow <- workflows::workflow() %>%
                 workflows::add_model(elastic_mod) %>%
                 workflows::add_recipe(RA_rec)
```

---

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
EN_grid <- dials::grid_regular(penalty(),
                               mixture(),
                               levels = 50)
```

---

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
EN_res <- elastic_wflow %>%
              tune::tune_grid(resamples = req_folds_spatial,
                              grid = EN_grid,
                              control = control_grid(verbose = TRUE, save_pred = TRUE),
                              metrics = metric_set(rmse))

#look at top 5 EN models
top_EN_models <- EN_res %>%
                    tune::show_best("rmse", n = 5)
top_EN_models

#default visualization
EN_auto_plot <- EN_res %>% autoplot()
EN_auto_plot

#save file
EN_auto_fig_file = here("results","EN-auto-plot.png")
ggsave(filename = EN_auto_fig_file, plot = EN_auto_plot)
```

---

### 5. Identify Best Model
```{r}
#select the EN model with the lowest rmse
EN_lowest_rmse <- EN_res %>%
                      tune::select_best("rmse")

#finalize the workflow by using the selected RF model
best_EN_wflow <- elastic_wflow %>%
                      tune::finalize_workflow(EN_lowest_rmse)
best_EN_wflow

#one last fit on the training data
best_EN_fit <- best_EN_wflow %>%
                    parsnip::fit(data = req_train_data)
```

---

### 6. Model evaluation
```{r}
#extract model from final fit
x_EN <- best_EN_fit$fit$fit$fit

#plot most important predictors in the model
EN_plot <- vip::vip(x_EN, 
                    num_features = 20,
                    aesthetics = list(fill = viridis::viridis(20, option = "D", begin = 0.9, end = 0)))
EN_plot

#save manually as "EN-plot.png" -- ggsave throws errors

#find predictions and intervals
EN_resid <- best_EN_fit %>%
                broom.mixed::augment(new_data = req_train_data) %>%
                dplyr::select(.pred, log_req_amt) %>%
                dplyr::mutate(.resid = .pred - log_req_amt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline is a 45 degree line, along which the results should fall
EN_pred_act <- ggplot2::ggplot(EN_resid, aes(x = .pred, y = log_req_amt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) +
                        geom_point(color = "#3b528b") +
                        scale_y_continuous(expand = c(0,0),
                                           limits = c(-30, 25)) +
                        scale_x_continuous(limits = c(0, 25)) +
                        labs(title = "EN Fit: Actual vs. Predicted Log(Requested Funds)",
                              x = "Predicted Log(Requested Funds) ($)",
                              y = "Actual Log(Requested Funds) ($)")
EN_pred_act

#save file
EN_pred_act_file = here("results","EN-pred-act.png")
ggsave(filename = EN_pred_act_file, plot = EN_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
EN_resids <- ggplot2::ggplot(EN_resid, aes(x = as.numeric(row.names(EN_resid)), y = .resid))+
                            geom_hline(yintercept = 0, color = "red", lty = 2) +
                            geom_point(color = "#3b528b") +
                            labs(title = "EN Fit: Residuals",
                                  x = "Observation Number",
                                  y = "Residual")
EN_resids

#save file
EN_resid_file = here("results","EN-resid.png")
ggsave(filename = EN_resid_file, plot = EN_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
EN_pred_resid <- ggplot2::ggplot(EN_resid, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point(color = "#3b528b") +
                          labs(title = "EN Fit: Residuals vs Fitted Log(Requested Funds)",
                                x = "Fitted Log(Requested Funds) ($)",
                                y = "Residual")
EN_pred_resid

#save file
EN_pred_resid_file = here("results","EN-pred-resid.png")
ggsave(filename = EN_pred_resid_file, plot = EN_pred_resid)

#print the 10 best performing hyperparameter sets
EN_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err)

#print the best model
EN_performance <- EN_res %>% tune::show_best(n = 1)
EN_performance

#compare model performance to null model (and other models)
EN_RMSE <- EN_res %>%
              tune::show_best(n = 1) %>%
              dplyr::transmute(
                rmse = round(mean, 3),
                SE = round(std_err, 4),
                model = "EN") %>%
             dplyr::bind_rows(RF_RMSE)
EN_RMSE
```
The Elastic Net model places great emphasis on FEMA Cost Share average above all other predictors. The next closest predictors include Hawaii, New York, New Mexico, and DC as key predictors. The EN model has an improved RMSE, but the null model technically still has a lower RMSE.

---

## LASSO Model

<br>

### 1. Model Specification
```{r}
#run parallels to determine number of cores
cores <- parallel::detectCores() - 1
cores

cl <- makeCluster(cores)

registerDoParallel(cl)

#define the LASSO model
#mixture = 1 identifies the model to be a LASSO model
lasso_mod <- linear_reg(mode = "regression",
                           penalty = tune(),
                           mixture = 1) %>%
                parsnip::set_engine("glmnet")
```

---

### 2. Workflow Definition
```{r}
#define lasso workflow
lasso_wflow <- workflows::workflow() %>%
                 workflows::add_model(lasso_mod) %>%
                 workflows::add_recipe(RA_rec)
```

---

### 3. Tuning Grid Specification
```{r}
#tuning grid specification
lasso_grid <- tibble(penalty = 10^seq(-3, 0, length.out = 30))

#5 lowest penalty values
lasso_grid %>%
  dplyr::top_n(-5)

#5 highest penalty values
lasso_grid %>%
  dplyr::top_n(5)
```

---

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
lasso_res <- lasso_wflow %>%
              tune::tune_grid(resamples = req_folds_spatial,
                              grid = lasso_grid,
                              control = control_grid(verbose = TRUE, save_pred = TRUE),
                              metrics = metric_set(rmse))

#look at top 5 EN models
top_lasso_models <- lasso_res %>%
                        tune::show_best("rmse", n = 5)
top_lasso_models

#default visualization
lasso_auto_plot <- lasso_res %>% autoplot()
lasso_auto_plot

#save file
lasso_auto_fig_file = here("results","lasso-auto-plot.png")
ggsave(filename = lasso_auto_fig_file, plot = lasso_auto_plot)
```

---

### 5. Identify Best Model
```{r}
#select the LASSO model with the lowest rmse
lasso_lowest_rmse <- lasso_res %>%
                        tune::select_best("rmse")

#finalize the workflow by using the selected lasso model
best_lasso_wflow <- lasso_wflow %>%
                      tune::finalize_workflow(lasso_lowest_rmse)
best_lasso_wflow

#one last fit on the training data
best_lasso_fit <- best_lasso_wflow %>%
                    parsnip::fit(data = req_train_data)
```

---

### 6. Model evaluation
```{r}
#extract model from final fit
x_lasso <- best_lasso_fit$fit$fit$fit

#plot most important predictors in the model
lasso_plot <- vip::vip(x_lasso, 
                    num_features = 20,
                    aesthetics = list(fill = viridis::viridis(20, option = "D", begin = 0.9, end = 0)))
lasso_plot

#save manually as "LASSO-plot.png" -- ggsave throws errors

#find predictions and intervals
lasso_resid <- best_lasso_fit %>%
                broom.mixed::augment(new_data = req_train_data) %>%
                dplyr::select(.pred, log_req_amt) %>%
                dplyr::mutate(.resid = .pred - log_req_amt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline is a 45 degree line, along which the results should fall
lasso_pred_act <- ggplot2::ggplot(lasso_resid, aes(x = .pred, y = log_req_amt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) +
                        geom_point(color = "#3b528b") +
                        scale_y_continuous(expand = c(0,0),
                                           limits = c(-30, 25)) +
                        scale_x_continuous(limits = c(0, 25)) +
                        labs(title = "LASSO Fit: Actual vs. Predicted Log(Requested Funds)",
                              x = "Predicted Log(Requested Funds) ($)",
                              y = "Actual Log(Requested Funds) ($)")
lasso_pred_act

#save file
lasso_pred_act_file = here("results","lasso-pred-act.png")
ggsave(filename = lasso_pred_act_file, plot = lasso_pred_act)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
lasso_resids <- ggplot2::ggplot(lasso_resid, aes(x = as.numeric(row.names(lasso_resid)), y = .resid))+
                                geom_hline(yintercept = 0, color = "red", lty = 2) +
                                geom_point(color = "#3b528b") +
                                labs(title = "LASSO Fit: Residuals",
                                      x = "Observation Number",
                                      y = "Residual")
lasso_resids

#save file
lasso_resid_file = here("results","lasso-resid.png")
ggsave(filename = lasso_resid_file, plot = lasso_resids)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
lasso_pred_resid <- ggplot2::ggplot(lasso_resid, aes(x = .pred, y = .resid))+
                                    geom_hline(yintercept = 0, color = "red", lty = 2) +
                                    geom_point(color = "#3b528b") +
                                    labs(title = "LASSO Fit: Residuals vs Fitted Log(Requested Funds)",
                                          x = "Fitted Log(Requested Funds) ($)",
                                          y = "Residual")
lasso_pred_resid

#save file
lasso_pred_resid_file = here("results","lasso-pred-resid.png")
ggsave(filename = lasso_pred_resid_file, plot = lasso_pred_resid)

#print the 10 best performing hyperparameter sets
lasso_res %>%
  tune::show_best(n = 10) %>%
  dplyr::select(rmse = mean, std_err)

#print the best model
lasso_performance <- lasso_res %>% tune::show_best(n = 1)
lasso_performance

#compare model performance to null model (and other models)
lasso_RMSE <- lasso_res %>%
              tune::show_best(n = 1) %>%
              dplyr::transmute(
                rmse = round(mean, 3),
                SE = round(std_err, 4),
                model = "LASSO") %>%
             dplyr::bind_rows(EN_RMSE)
lasso_RMSE

#save combined table
model_comp_file = here("results", "ML-model-comp.Rds")
saveRDS(lasso_RMSE, file = model_comp_file)
```
The LASSO model fit is almost identical to the Elastic Net model fit.

---

## Model Selection and Evaluation
Based on the results of the models above, the null model actually has the best fit, which warrants further investigation (Especially with the outliers). In the mean time, we'll fit the LASSO model as it has the best fit out of the models.
```{r}
#fit model to training set but evaluate on the test data
lasso_fit_test <- best_lasso_wflow %>%
                    tune::last_fit(split = req_data_split)

#compare test performance against training performance
lasso_rmse_test <- collect_metrics(lasso_fit_test) %>%
                   dplyr::select(rmse = .estimate) %>%
                   dplyr::mutate(data = "test")

lasso_comp <- lasso_RMSE %>%
                  dplyr::filter(model == "LASSO") %>%
                  dplyr::transmute(rmse, data = "train") %>%
                  dplyr::bind_rows(lasso_rmse_test) %>%
                  dplyr::slice(-3) #don't know why the third row shows up
lasso_comp
#RMSEs are different but not as much as previous iterations

#save combined table
test_comp_file = here("results", "test-train-model-comp.Rds")
saveRDS(lasso_comp, file = test_comp_file)

#find predictions and intervals
lasso_resid_fit <- lasso_fit_test %>%
                      broom.mixed::augment() %>%
                      dplyr::select(.pred, log_req_amt) %>%
                      dplyr::mutate(.resid = .pred - log_req_amt)

#plot model predictions from tuned model versus actual outcomes
#geom_abline draws a 45 degree line, along which the results should fall
lasso_pred_act_test <- ggplot2::ggplot(lasso_resid_fit, aes(x = .pred, y = log_req_amt)) +
                        geom_abline(slope = 1, intercept = 0, color = "red", lty = 2) + 
                        geom_point(color = "#3b528b") +
                        labs(title = "LASSO Test Fit: Predicted vs. Actual Log(Requested Funds)",
                              x = "Predicted Log(Requested Funds) ($)",
                              y = "Actual Log(Requested Funds) ($)")
lasso_pred_act_test

#save file
lasso_pred_act_test_file = here("results","lasso-pred-act-test.png")
ggsave(filename = lasso_pred_act_test_file, plot = lasso_pred_act_test)

#plot model with residuals
#the geom_hline plots a straight horizontal line along which the results should fall
lasso_resids_test <- ggplot2::ggplot(lasso_resid_fit, aes(x = as.numeric(row.names(lasso_resid_fit)), y = .resid))+
                      geom_hline(yintercept = 0, color = "red", lty = 2) +
                      geom_point(color = "#3b528b") +
                      labs(title = "LASSO Test Fit: Residuals",
                            x = "Observation Number",
                            y = "Residual")
lasso_resids_test

#save file
lasso_resid_test_file = here("results","lasso-resid-test.png")
ggsave(filename = lasso_resid_test_file, plot = lasso_resids_test)

#plot model fit vs residuals
#the geom_hline plots a straight horizontal line along which the results fall
lasso_pred_resid_test <- ggplot2::ggplot(lasso_resid_fit, aes(x = .pred, y = .resid))+
                          geom_hline(yintercept = 0, color = "red", lty = 2) +
                          geom_point(color = "#3b528b") +
                          labs(title = "LASSO Test Fit: Residuals vs Fitted Log(Requested Funds)",
                                x = "Predicted Log(Requested Funds) ($)",
                                y = "Residuals")
lasso_pred_resid_test

#save file
lasso_pred_resid_test_file = here("results","lasso-pred-resid-test.png")
ggsave(filename = lasso_pred_resid_test_file, plot = lasso_pred_resid_test)
```