### 2021-12-13 code/analysis_code/analysis-linearmodels.Rmd

---
title: "Analysis: Linear Models"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---

---

## Introduction
This markdown imports the processed analysis data and uses the tidymodels framework to fit linear models. Follow the previous markdowns to generate the data.

<br>

This analysis examines the predictors of allocation of FEMA funds during disaster declarations. The following definitions exist for the analysis:

* The two outcomes of interest are Requested Amount (`ReqAmt`) and Obligated Amount (`OblAmt`).
* The primary predictor of interest is Incident Duration (`IncidentDuration`)
* Whichever outcome is not currently fitted to the model will *not* be considered a predictor of interest
* The outcomes of interest are log-transformed

<br>

For each outcome of interest, the following steps will be completed:

1. Fit a simple linear regression with the primary predictor of interest
2. Fit a multivariable linear regression
3. Comparison of the two models

---

## Required Packages
The following R packages are required for this script:

* here: for path setting
* tidyverse: for all packages in the Tidyverse (ggplot2, dyplr, tidyr, readr, purr, tibble, stringr, forcats)
* skimr: for data summarization
* tidymodels: for data modeling

```{r libraries, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load required packages
library(here) #to set paths
library(tidyverse) #for data processing
library(skimr) #for data summarization
library(tidymodels) #for data modeling

#global environment options
# formatting for script to avoid scientific notation output
options(scipen=999)
```

---

## Load Processed Data
Load the data generated from the markdowns in the `processing_code` folder.
```{r load data}
#define file path to req data
data_location_req <- here::here("data", "processed_data", "analysisdata-RA.rds")

#load req data
analysis_data_req <- readRDS(data_location_req)

#define file path to obl data
data_location_obl <- here::here("data", "processed_data", "analysisdata-OA.rds")

#load obl data
analysis_data_obl <- readRDS(data_location_obl)
```

---

## Data Overview
To better understand the data, let's use `summarytools` to better visualize the data.
```{r data overview}
skimr::skim(analysis_data_req)

skimr::skim(analysis_data_obl)
```

---

## Data Filtering
We first need to address the NA values, and then we will make two datasets with the only the relevant variables for each:

* state
* IncidentYear
* IncidentMonth
* declarationType
* incidentType
* IncidentDuration
* ResponseDays
* IH
* PA
* HM
* Population
* Counties
* TotalAgencies
* FEMACSAvg

```{r data subsets}
#log transform ReqAmt
analysis_data_req$logReqAmt <- log(analysis_data_req$ReqAmt)

#drop unnecessary variables
req_data <- analysis_data_req %>%
              dplyr::select(-c(ReqAmt, disasterNumber, declarationTitle))

#log transform OblAmt
analysis_data_obl$logOblAmt <- log(analysis_data_obl$OblAmt)

#drop unnecessary variables
obl_data <- analysis_data_obl %>%
              dplyr::select(-c(OblAmt, disasterNumber, declarationTitle))
```

---

## Outcome of Interest: Requested FEMA Funds
We will start with requested FEMA funds as the outcome of interest.

<br>

### Data Set Up
Split the data randomly into train and test datasets, and define cross validation
```{r req data splitting}
#fix random numbers by setting the seed (helps reproducibility)
set.seed(123)

#put 3/4 of data into the training set
req_data_split <- rsample::initial_split(req_data, prop = 3/4)

#create dataframes for the two sets
req_train_data <- rsample::training(req_data_split)
req_test_data <- rsample::testing(req_data_split)
```

---

### Simple Linear Regression
Create the SLR using the `tidymodels` setup.
```{r req SLR}
#define lr mod
lr_mod <- parsnip::linear_reg() %>%
          parsnip::set_engine("lm")

#define recipe
req_SLR_rec <- recipes::recipe(logReqAmt ~ IncidentDuration, data = req_train_data)

#define SLR workflow
req_SLR_wflow <- workflows::workflow() %>%
                   workflows::add_model(lr_mod) %>%
                   workflows::add_recipe(req_SLR_rec)

#fit the training data to the workflow
req_SLR_fit <- req_SLR_wflow %>%
                  parsnip::fit(data = req_train_data)

#create tibble for model fit using broom and extract
req_SLR_fit %>%
  workflowsets::extract_fit_parsnip() %>%
  broom::tidy()
```

<br>

Create box and whisker plot for the fit output.
```{r req SLR bp}
req_SLR_bp <- broom.mixed::tidy(req_SLR_fit) %>%
                dotwhisker::dwplot(dot_args = list(size = 2, color = "blue"),
                                   whisker_args = list(color = "blue"),
                                   vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
req_SLR_bp

#save image
req_SLR_bp_file = here("results","req_SLR_bp.png")
ggsave(filename = req_SLR_bp_file, plot = req_SLR_bp)
```

<br>

Predict requested funds from FEMA  based on the model above for the training data set.
```{r req SLR predict}
#use predict to predict funds
stats::predict(req_SLR_fit, req_train_data)

#create df with model predictions and actual measures
req_SLR_aug <- augment(req_SLR_fit, req_train_data)

#add labeling variables for later summarization
req_SLR_train <- req_SLR_aug %>%
                    dplyr::mutate(data = "train",
                                  model = "SLR")
```

<br>

Assess goodness of fit measures.
```{r req SLR eval}
#using glace
modelsummary::glance(req_SLR_fit)

#using the prediction capacity
req_SLR_train %>% yardstick::metrics(truth = logReqAmt, estimate = .pred)
```
This isn't a great model. Time to try adding more predictors.

---

### Multivariable Linear Regression
Create the MVR using the `tidymodels` setup.
```{r req MVR}
#define linear mod
lr_mod <- linear_reg() %>%
              set_engine("lm")

#define recipe
req_MVR_rec <- recipes::recipe(logReqAmt ~ ., data = req_train_data)

#define SLR workflow
req_MVR_wflow <- workflows::workflow() %>%
                   workflows::add_model(lr_mod) %>%
                   workflows::add_recipe(req_MVR_rec)

#fit the training data to the workflow
req_MVR_fit <- req_MVR_wflow %>%
                  parsnip::fit(data = req_train_data)

#create tibble for model fit using broom and extract
req_MVR_fit %>%
  workflowsets::extract_fit_parsnip() %>%
  broom::tidy()
```

<br>

Create box and whisker plot for the fit output.
```{r req MVR bp}
#basic box and whisker plot
req_MVR_bp <- broom.mixed::tidy(req_MVR_fit) %>%
                dotwhisker::dwplot(dot_args = list(size = 2, color = "blue"),
                                   whisker_args = list(color = "blue"),
                                   vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
req_MVR_bp

#there's a lot of information here, but hard to identify the ones that are significant due to volume

#box and whisker plot for significant predictors
#first filter significant results
req_MVR_fit_sig <- broom.mixed::tidy(req_MVR_fit) %>%
                      dplyr::filter(p.value < 0.05)

#box and whisker plot for lm_fit2 significant predictors
req_MVR_fit_bp2 <- req_MVR_fit_sig %>%
                        dotwhisker::dwplot(dot_args = list(size = 2, color = "blue"),
                                           whisker_args = list(color = "blue"),
                                           vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
req_MVR_fit_bp2

#save file
req_MVR_bp_file = here("results","req_MVR_bp.png")
ggsave(filename = req_MVR_bp_file, plot = req_MVR_fit_bp2)
```
This model identifies total number of federal agencies, the state/territory being Puerto Rico, and emergency declarations as significant predictors of requested FEMA funding. It makes sense that an emergency declaration is a protective factor, as it is a smaller event than a major disaster declaration.

<br>

Predict requested funds from FEMA  based on the model above for the training data set.
```{r req MVR predict}
#use predict to predict funds
stats::predict(req_MVR_fit, req_train_data)

#create df with model predictions and actual measures
req_MVR_aug <- augment(req_MVR_fit, req_train_data)

#add labeling variables for later summarization
req_MVR_train <- req_MVR_aug %>%
                    dplyr::mutate(data = "train",
                                  model = "MVR")
```

<br>

Assess goodness of fit measures.
```{r req MVR eval}
#using glace
modelsummary::glance(req_MVR_fit)

#using the prediction capacity
req_MVR_train %>% yardstick::metrics(truth = logReqAmt, estimate = .pred)
```
While this is an improvement, it's still not a great model. Let's compare the two side-by-side using RMSE.

---

### Requested Funds: Model Evaluation
Formally compare the two linear models
```{r req lm eval}
#First create comprehensive df with all of the predictions
req_lm_preds <- bind_rows(req_MVR_train,
                          req_SLR_train)

#Use the yardstick package to calculate the metrics on resamples
req_lm_metrics <- req_lm_preds %>%
                      dplyr::group_by(data, model) %>%
                      yardstick::metrics(truth = logReqAmt, estimate = .pred)

#for interpretation: rmse = Root Mean Squared Error (RMSE), rsq = R^2, mae = Mean Absolute Error (MAE)
#make a df that displays RMSE and R^2 for each model and data (all dplyr functions)
req_lm_RMSE <- req_lm_metrics %>%
                    filter(.metric == "rmse" | .metric == "rsq") %>%
                    select(-c(.estimator, data))

#save comparison table
req_lm_eval_file = here("results", "lm-req-lm-eval.Rds")
saveRDS(req_lm_RMSE, file = req_lm_eval_file)
```
In comparing the two models, the full model certainly improves the fit, but neither is ideal. Alternative models should be considered.

---

## Outcome of Interest: Obligated FEMA Funds
We will now examine obligated FEMA funds as the outcome of interest.

<br>

### Data Set Up
Split the data randomly into train and test datasets, and define cross validation
```{r obl data splitting}
#fix random numbers by setting the seed (helps reproducibility)
set.seed(123)

#put 3/4 of data into the training set
obl_data_split <- rsample::initial_split(obl_data, prop = 3/4)

#create dataframes for the two sets
obl_train_data <- rsample::training(obl_data_split)
obl_test_data <- rsample::testing(obl_data_split)
```

---

### Simple Linear Regression
Create the SLR using the `tidymodels` setup.
```{r obl SLR}
#define lr mod
lr_mod <- parsnip::linear_reg() %>%
          parsnip::set_engine("lm")

#define recipe
obl_SLR_rec <- recipes::recipe(logOblAmt ~ IncidentDuration, data = obl_train_data)

#define SLR workflow
obl_SLR_wflow <- workflows::workflow() %>%
                   workflows::add_model(lr_mod) %>%
                   workflows::add_recipe(obl_SLR_rec)

#fit the training data to the workflow
obl_SLR_fit <- obl_SLR_wflow %>%
                  parsnip::fit(data = obl_train_data)

#create tibble for model fit using broom and extract
obl_SLR_fit %>%
  workflowsets::extract_fit_parsnip() %>%
  broom::tidy()
```

<br>

Create box and whisker plot for the fit output.
```{r obl SLR bp}
obl_SLR_bp <- broom.mixed::tidy(obl_SLR_fit) %>%
                dotwhisker::dwplot(dot_args = list(size = 2, color = "blue"),
                                   whisker_args = list(color = "blue"),
                                   vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
obl_SLR_bp

#save image
obl_SLR_bp_file = here("results","obl_SLR_bp.png")
ggsave(filename = obl_SLR_bp_file, plot = obl_SLR_bp)
```

<br>

Predict requested funds from FEMA  based on the model above for the training data set.
```{r obl SLR predict}
#use predict to predict funds
stats::predict(obl_SLR_fit, obl_train_data)

#create df with model predictions and actual measures
obl_SLR_aug <- augment(obl_SLR_fit, obl_train_data)

#add labeling variables for later summarization
obl_SLR_train <- obl_SLR_aug %>%
                    dplyr::mutate(data = "train",
                                  model = "SLR")
```

<br>

Assess goodness of fit measures.
```{r obl SLR eval}
#using glace
modelsummary::glance(obl_SLR_fit)

#using the prediction capacity
obl_SLR_train %>% yardstick::metrics(truth = logOblAmt, estimate = .pred)
```
This isn't a great model. Time to try adding more predictors.

---

### Multivariable Linear Regression
Create the MVR using the `tidymodels` setup.
```{r obl MVR}
#define linear mod
lr_mod <- linear_reg() %>%
              set_engine("lm")

#define recipe
obl_MVR_rec <- recipes::recipe(logOblAmt ~ ., data = obl_train_data)

#define SLR workflow
obl_MVR_wflow <- workflows::workflow() %>%
                   workflows::add_model(lr_mod) %>%
                   workflows::add_recipe(obl_MVR_rec)

#fit the training data to the workflow
obl_MVR_fit <- obl_MVR_wflow %>%
                  parsnip::fit(data = obl_train_data)

#create tibble for model fit using broom and extract
obl_MVR_fit %>%
  workflowsets::extract_fit_parsnip() %>%
  broom::tidy()
```

<br>

Create box and whisker plot for the fit output.
```{r obl MVR bp}
#basic box and whisker plot
obl_MVR_bp <- broom.mixed::tidy(obl_MVR_fit) %>%
                dotwhisker::dwplot(dot_args = list(size = 2, color = "blue"),
                                   whisker_args = list(color = "blue"),
                                   vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
obl_MVR_bp

#there's a lot of information here, but hard to identify the ones that are significant due to volume

#box and whisker plot for significant predictors
#first filter significant results
obl_MVR_fit_sig <- broom.mixed::tidy(obl_MVR_fit) %>%
                      dplyr::filter(p.value < 0.05)

#box and whisker plot for lm_fit2 significant predictors
obl_MVR_fit_bp2 <- obl_MVR_fit_sig %>%
                        dotwhisker::dwplot(dot_args = list(size = 2, color = "blue"),
                                           whisker_args = list(color = "blue"),
                                           vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
obl_MVR_fit_bp2

#save file
obl_MVR_bp_file = here("results","obl_MVR_bp.png")
ggsave(filename = obl_MVR_bp_file, plot = obl_MVR_fit_bp2)
```
This model identifies total number of federal agencies, the state/territory being Puerto Rico, and emergency declarations as significant predictors of obligated FEMA funding. It makes sense that an emergency declaration is a protective factor, as it is a smaller event than a major disaster declaration.

<br>

Predict obligated funds from FEMA  based on the model above for the training data set.
```{r obl MVR predict}
#use predict to predict funds
stats::predict(obl_MVR_fit, obl_train_data)

#create df with model predictions and actual measures
obl_MVR_aug <- augment(obl_MVR_fit, obl_train_data)

#add labeling variables for later summarization
obl_MVR_train <- obl_MVR_aug %>%
                    dplyr::mutate(data = "train",
                                  model = "MVR")
```

<br>

Assess goodness of fit measures.
```{r obl MVR eval}
#using glace
modelsummary::glance(obl_MVR_fit)

#using the prediction capacity
obl_MVR_train %>% yardstick::metrics(truth = logOblAmt, estimate = .pred)
```
While this is an improvement, it's still not a great model. Let's compare the two side-by-side using RMSE.

---

### Obligated Funds: Model Evaluation
Formally compare the two linear models
```{r obl lm eval}
#First create comprehensive df with all of the predictions
obl_lm_preds <- bind_rows(obl_MVR_train,
                          obl_SLR_train)

#Use the yardstick package to calculate the metrics on resamples
obl_lm_metrics <- obl_lm_preds %>%
                      dplyr::group_by(data, model) %>%
                      yardstick::metrics(truth = logOblAmt, estimate = .pred)

#for interpretation: rmse = Root Mean Squared Error (RMSE), rsq = R^2, mae = Mean Absolute Error (MAE)
#make a df that displays RMSE and R^2 for each model and data (all dplyr functions)
obl_lm_RMSE <- obl_lm_metrics %>%
                    filter(.metric == "rmse" | .metric == "rsq") %>%
                    select(-c(.estimator, data))

#save comparison table
obl_lm_eval_file = here("results", "lm-obl-lm-eval.Rds")
saveRDS(obl_lm_RMSE, file = obl_lm_eval_file)
```
In comparing the two models, the full model certainly improves the fit, but neither is ideal. Alternative models should be considered.

---

## Compare Requested and Obligated Models
As one last step for the linear models, let's compare the metrics from the obligated and requested funds models.
```{r lm comp}
#add outcome labels
req_lm_RMSE$outcome <- "Requested"
obl_lm_RMSE$outcome <- "Obligated"

#combine into one table
lm_comp <- dplyr::bind_rows(req_lm_RMSE,
                            obl_lm_RMSE)

#clean up the table
lm_comp$label <- paste(lm_comp$model, lm_comp$.metric, sep=": ")

lm_comp_tbl <- lm_comp %>%
                  subset(select = -c(1:2))

lm_comp_table <- xtabs(.estimate ~ label + outcome, lm_comp_tbl)
lm_comp_table

#save file
lm_eval_file = here("results", "lm-eval.Rds")
saveRDS(lm_comp_table, file = lm_eval_file)
```
The two outcomes yield virtually identical models.